Differential spatial regulation and activation of integrin nanoclusters inside focal adhesions

Version 1

Abstract

Abstract

α5β1 and αvβ3 integrins are core components of focal adhesions (FAs) involved in cell attachment, migration and mechanobiology-dependent processes. Recent works indicate that both integrins organize in nanoclusters inside FAs, with sub-populations of active and inactive β1 nanoclusters. However, whether both integrins work in concert or their activities are spatially regulated is not fully understood. Using dual-color super-resolution microscopy we show that integrins α5β1 and αvβ3 exhibit similar functional nanoscale segregation. Notably, both integrins never mix at the nanoscale indicating that their functions might be spatially regulated. We find a nearly 1:1 stoichiometry between active integrin and adaptor nanoclusters suggesting that coordinated integrin activation occurs via the concurrent engagement of adaptor nanoclusters. Interestingly, α5β1 nanoclusters preferentially localize at the FA periphery in close proximity to adaptor nanoclusters, establishing regions of multi-nanocluster enrichment, whereas αvβ3 nanoclusters uniformly distribute throughout FAs. Overall, our results show that adhesion proteins arrange as modular nanoscale units that distinctively organize inside FAs to spatially regulate integrin activation and function.

Introduction

Adhesion complexes (ACs) in adherent cells, such as focal adhesions (FAs) and fibrillar adhesions (fAs) are places for cell attachment, mechanosensing and mechanotransduction (1–4). These structures are composed by different integrin receptors that bind to their ligands on the extracellular matrix (ECM), and by a myriad of intracellular adaptors, proteins that dynamically connect integrins to the actin cytoskeleton machinery relaying the transmission of mechanical forces in a bidirectional manner (5–7). Studies over the last two decades have provided a wealth of information focused on understanding these complexes, the proteins they contain, their role in cellular processes, such as mechanosensing, migration and proliferation, as well as how influential their regulation is in developmental and pathogenic processes (8–15).

Integrins α5β1 and αvβ3 are transmembrane heterodimeric proteins critical for the formation of ACs. Both α5β1 and αvβ3 bind to the RGD peptide motif of certain ECM components, such as fibronectin (16–18), and are linked to the cell’s cytoskeleton via many of the same adaptor proteins (19–22). The apparent redundancy of different integrins binding to the same ligand and same adaptor proteins is intriguing and has been the subject of intense research over the past years. Earlier studies on primary fibroblasts showed that these integrins segregate in different adhesion structures, indicating different roles, with α5β1 being enriched in fAs (23), adhesive sites at central regions of the cell, and αvβ3 being confined to focal complexes at the base of cellular protrusions, and to FAs, which anchor contractile stress fibers at the periphery of the cell (24). However, more recent super-resolution microscopy studies showed that both types of integrins are simultaneously present inside FAs (25). Within FAs, α5β1 has been proposed as the integrin responsible for adhesion strength while αvβ3 enables mechanotransduction, as cells exposed to α5β1 blocking antibody exhibited weaker attachment than those whose αvβ3 integrins were blocked (5). These results are also aligned to in vitro reconstitution experiments, that have assessed the catch-bond behavior of both integrins with fibronectin. In this type of bond, the lifetime of the bond increases as the applied force increases, up to a certain threshold, beyond which the bond ruptures. The rupture force for α5β1 binding to its ligand has been found to be higher (∼10–30 pN) than that of αvβ3 (∼5–25 pN) (26, 27). Nevertheless, single molecule tracking data indicated that αvβ3 integrins are mainly stationary inside FAs and more stably bound to the ligand, favoring adhesion as its main role (25). It has been also shown that α5β1 is involved in Rho activation during late spreading, enhancing RhoA-mediated actomyosin contractility (24, 28), supporting the idea that α5β1 is involved in force-mediated adhesion reinforcement. In a tour-de-force effort, Fässler and colleagues found that only the simultaneous expression of both integrins enabled cells to sense the rigidity of the ECM and to respond by reinforcing the actomyosin structure (29). These results, supported by others (30), have led to the recent notion that both integrins work in a synergistic manner, exhibiting cooperativity during cell adhesion, migration and mechanosensing, although the mechanisms are far from being fully resolved (30).

The highly dynamic conformational states exhibited by integrins together with their lateral and axial nanoscale organization might play important roles regulating the differential or synergistic functions of different integrins inside FAs. Indeed, it has been shown that integrins can rapidly transition between different conformational states that relate to their activation and binding capacity to their ligands (25, 31, 32). Moreover, in addition to the nanoscale axial segregation of different AC components inside FAs (33), it has been reported that integrins as well as different adaptor proteins also segregate laterally into small nanoclusters inside FAs (34–37). Recent super-resolution studies have further shown that β1 integrins segregate in both active and inactive nanoclusters inside FAs (36) and single-molecule tension measurements uncovered distinct sub-populations of load bearing integrins and a high fraction of unengaged integrins (38, 39). Together, these results suggest that at any given time only a fraction of integrins might be active inside FAs and engaged with their adaptor partners. Such a dynamic scenario would allow integrins to differentially interact with their environment and possibly diversify their functions. As α5β1 and αvβ3 have different activation states and are both contained in FAs, it is plausible that their spatial regulation and activation state inside FAs provides a physical mechanism to tune their functions. However, such studies have not been reported yet.

Here, we combined dual-color super-resolution microscopy together with robust analytical tools to study a number of endogenous FA proteins on human fibroblast cells. Specifically, we imaged one of the two integrins (α5β1/αvβ3), their full population or active conformations, with each of three key adaptors proteins, paxillin, talin and vinculin. Our results show that each of these proteins form segregated nanoclusters of remarkably similar size and molecular density inside FAs, and interestingly, also on the membrane outside of ACs. Moreover, we find sub-populations of active and inactive β1 and β3 nanoclusters inside FAs, which spatially segregate from each other and provide evidence for a nearly 1:1 stoichiometry between talin, vinculin and active integrin nanoclusters. Interestingly, we find that αvβ3 nanoclusters organize randomly inside FAs regardless of their activation state, whereas α5β1 and in particular the active conformation of β1 nanoclusters preferentially organize at the edges of FAs. Overall, we propose that the nano- and meso-scale organization of α5β1 and αvβ3 integrins and their distinct lateral segregation and activation state are key factors in the differentiation of their roles within FAs.

Results

Proteins of the ACs organize as universal nanoclusters across regions of the basal membrane

We employed dual-color super-resolution microscopy based on stochastic optical reconstruction microscopy (STORM) to determine how adhesion proteins (integrins and key adaptor proteins) are spatially organized at the nanoscale. Human fibroblasts (HFF-1 cells) were seeded on fibronectin (FN)-coated glass slides for 24 h, after which the cells were processed for immunofluorescence. One integrin (either the total population of α5β1 or αvβ3, or the active population specific to the β1 or β3 integrin subunit) was labeled together with one adaptor protein (paxillin, C terminus of talin-1, or vinculin) and then imaged using STORM. We note that while the pan antibodies used to target the full population of integrins are specific to α5β1 or αvβ3, (with the anti-α5 antibody binding the α5 subunit, which only dimerizes with β1, and the anti-αvβ3 targeting this specific heterodimer), the active antibodies target the β subunits. These β subunits can dimerize with other α subunits, but with our experimental conditions, where FN is the most abundant ligand (see Materials & Methods), the primary integrins being probed are the heterodimers, α5β1 or αvβ3. Given the huge number of adaptor proteins that have been reported to interact with integrins (19, 20), we chose for paxillin, talin and vinculin since they belong to the core-components of the adhesome (21) and crucially locate at distinct axial layers of functional activity, i.e., paxillin at the signaling layer, while talin and vinculin locate at the force-transduction layer (33).

Super-resolution images qualitatively revealed an inhomogeneous distribution for all the proteins examined, forming small and laterally segregated nanoclusters inside FAs (Fig. 1A– D). To quantitatively substantiate these observations, we first used a mask to classify all the single molecule localization events according to their location on the membrane, i.e., FAs, fAs or outside ACs (see Materials & Methods, Fig. S1A-F). To group localizations into nanoclusters we used a density-based spatial clustering of applications with noise (DBSCAN) algorithm (40), which has been extensively used for quantifying protein nanoclustering from single molecule localization images, including β1 integrins inside FAs (36, 41). Since STORM data consist of lists of single-molecule localizations and each labeled protein can be localized multiple times due to fluorophore blinking (42), it is important to first assess the number of localizations detected due to the multiple blinking events of a single fluorophore. For this, we followed an approach similar to that reported by Spiess et al. to generate distributions of the number of localizations obtained from isolated fluorescent spots on glass next to the cells, for both channels, these will mostly correspond to individual antibodies (Abs) (Fig. 1E, Materials & Methods). By doing this approach one can determine the number of localizations corresponding to individual Abs and infer whether the clustered localizations observed correspond to individual or multiple proteins. The median values of the distributions were 5 localizations for Alexa 405-Alexa 647-conjugated Abs (Channel 1) and 3 localizations for Cy3–Alexa 647-conjugated Abs (Channel 2). Moreover, to further minimize the effect of fluorophore blinking on cluster identification, we imposed a threshold of 10 localizations as commonly used in the field (36, 41). Thus, any cluster identified by DBSCAN with 10 or more localizations was considered as a cluster of more than one protein (see Materials & Methods). DBSCAN classified between 70% to 80% of the localizations of each of the proteins as belonging to nanoclusters, regardless of whether they were located inside FAs or on regions of the membrane outside ACs (Fig. S1G-I). Notably, the number of localizations per nanocluster and their size were similar across the different adhesion proteins (integrins and adaptors) inside FAs, with a median of ∼20 localizations per nanocluster, likely corresponding to a few copies of the protein, and a diameter of ∼ 40–45 nm (Fig. 1F, G).

Proteins of FAs organize as universal nanoclusters of similar size and containing a similar number of proteins.

(A) Representative dual-color STORM image of an HFF-1 cell fixed after 24 h on FN and labeled for α5β1 (magenta) and paxillin (green). Scale bar: 5 μm. (B) Representative images of all the adhesions identified for α5β1 with paxillin (left), talin (middle) and vinculin (right), with zoomed-in regions shown below each panel. Scale bar top row: 5 μm, bottom row: 1 μm. (C) Representative images of all the adhesions identified for αvβ3 with paxillin (left), talin (middle) and vinculin (right), with zoomed-in regions shown below each panel. Scale bar top row: 5 μm, bottom row: 1 μm. (D) Representative images all the adhesions identified for of active β1 (act β1) with paxillin (left), active β3 (act β3) with paxillin (right), with zoomed-in regions shown below each panel. Scale bar top row: 5 μm, bottom row: 1 μm. (E) Distribution for the number of localizations per secondary antibody on glass next to cell regions for both of the secondary antibodies used in this study, goat anti-mouse Alexa 405–Alexa 647 (Ch1) and goat anti-rabbit Cy3–Alexa 647 (Ch2). The median number of localizations per spot corresponds to: 5 for Alexa 405-Alexa 647-conjugated antibodies and 3 for Cy3–Alexa 647-conjugated antibodies (see Materials & Methods). (F, G) Box-and-whisker plots showing the number of localizations per nanocluster (F) and the nanocluster diameter (G) for the proteins in this study. The box represents the range from the 25th to the 75th percentiles and whiskers from 10th to 90th percentiles, with the median indicated by a horizontal line. Points outside these whiskers correspond to outliers. Individual points (points within the whiskers not shown) correspond to the median value over all nanoclusters for each cell (for N values, i.e., number of independent experiments, see SI Table S5). No statistical differences were found using the multi comparison one-way ANOVA test, where not significant (ns) corresponds to p>0.05.

Similar results in terms of size and number of localizations per nanocluster were obtained when performing the analysis outside ACs (Fig. S2). In addition, the nanocluster characteristics were maintained when specifically labeling the active integrin (β1 or β3) populations (Fig. 1 D, F, G). These results are in good quantitative agreement with previous reports using STORM showing that both active and inactive integrin β1 nanoclusters have a median size of 40 nm and 35 nm, respectively, as well as comparable number of single molecule localizations per nanocluster (36). Overall, our results agree with previous work on α5β1 and provides new insight on the integrin αvβ3 and the main adaptor proteins, paxillin, talin and vinculin, indicating that organization into laterally segregated nanoclusters of similar size and number of molecules is maintained across different protein types. Interestingly, such nanoscale organization is independent of whether these adhesion proteins partition inside FAs or outside ACs (Fig. S2).

Global mesoscale organization of adhesion protein nanoclusters inside FAs

We next sought to investigate how the different adhesion proteins laterally organize inside FAs beyond their nanoclustering, i.e., at the mesoscale. For this, we first quantified the density of nanoclusters per unit area of the membrane, within FAs (Fig. 2A) and outside ACs (Fig. 2B). As expected, the density of nanoclusters was larger in FAs as compared to outside adhesions (∼5–10-fold increase) (Fig. 2A, B). Adaptor proteins, such as talin or vinculin, are shared between α5β1 and αvβ3 integrins (21), therefore we compared the density of the full population of integrin nanoclusters (∼36 µm−2 in FAs) to the density of adaptor protein nanoclusters (∼10– 12 µm−2 for talin and vinculin, ∼20 µm−2 for paxillin in FAs). These data reveal a significantly higher (∼2–4 fold) density of integrin nanoclusters in comparison to their adaptors and, consequently, a stoichiometric nanocluster mismatch between the integrins and the adaptors, regardless of whether they partition inside or outside FAs (Fig 2A, B). Since the active integrin conformation is the one engaged with the adaptors (43, 44), we also quantified the density of active integrin nanoclusters (active β1 and active β3) per unit area inside FAs. Our results revealed that only ∼ 30% of the total α5β1 nanocluster population were in an active conformation, and this percentage was even lower for β3 nanoclusters (∼ 20%) (Fig. 2A), indicating that at any given time there is a large percentage of inactive integrin nanoclusters inside FAs. We note that these percentages might be even lower considering that our Abs for mapping the active conformations are not exclusive for α5β1 and αvβ3, but rather to the total active β1 and β3 sub-populations. Notably, the total density of active β1 and β3 integrin nanoclusters stoichiometrically matches that of talin and vinculin nanoclusters, and it is only slightly outnumbered by paxillin nanoclusters inside FAs (Fig. 2A), suggesting that these active integrins might be fully engaged to their adaptors. The modest percentage of active β1 integrin nanoclusters observed in our data (∼ 30%, Fig. 2A) is in close agreement with a recent report from Spiess et al, obtained using a different super-resolution technique (36) and consistent with the high percentage of low-force bearing integrins measured by using single-molecule force sensors (39). In addition to the α5β1 results, our data also reveal the existence of a sub-population of active β3 nanoclusters inside FAs (Fig. 2A). Together, these results strengthen the hypothesis proposed by Spiess et al. of a coordinated mechanism that laterally segregates active and inactive nanoclusters inside FAs (36).

Mesoscale organization of integrin nanoclusters and adaptor proteins inside FAs.

(A, B) Mean density of nanoclusters per unit area inside FAs (A) and on the cell membrane outside adhesions (B), with α5β1 and actβ1 in light and dark green, respectively; and αvβ3 and actβ3 in light and dark blue, respectively. The densities of integrins nanoclusters per unit area are shown as stacked graphs to allow comparison between the total number of integrin nanoclusters (α5β1 and αvβ3) and the adaptor proteins. The bars show the mean ± st. dev of the distribution. Statistical significance between data of all proteins was tested using a multi comparison one-way ANOVA, assuming Gaussian distribution of points. The comparison between α5β1 and αvβ3, (actβ1 and actβ3) was done with an unpaired t-test, where ns = not significant, p>0.05; *, p<0.05; **, p<0.01; ***, p<0.001. The insets show representative images of nanoclusters for α5β1 (magenta) and paxillin (green) in FAs (A) and outside adhesions (B). Scale bar: 1 μm. (C) Representative dual-color STORM image of the total populations of α5β1 (magenta) and αvβ3 (green) integrins inside FAs. The panel on the far-right side corresponds to a zoomed-in region of a single FA (white box) showing spatially segregated α5β1 and αvβ3 nanoclusters. Scale bars: 2 μm, and 500 nm for the zoomed-in region. (D) Histogram of the EED-NN values between α5β1 and the different adaptors (green) and, between αvβ3 and the adaptors (blue). The black line in each plot corresponds to the histogram for the simulated data sets. The dashed red line at zero corresponds to nanoclusters contiguous to each other. Percentage of overlapping nanoclusters: α5β1-paxillin: exp = 28.4%, sim = 24.4%; α5β1-talin: exp = 6.7%, sim = 8.1%; α5β1-vinculin: exp = 6.0%, sim = 7.7%; αvβ3 -paxillin: exp =31.7%, sim=29.2%; αvβ3 -talin: exp =15.7%, sim=16.5; αvβ3-vinculin: exp =14.9%, sim=17.4%. (E) Histogram of the NND values between nanoclusters of the same protein. The black line in each plot corresponds to the histogram for the simulated data sets. Bin width is 10 nm.

Interestingly, despite this common feature, we found a significantly higher density of α5β1 nanoclusters as compared to αvβ3 (Fig. 2A, B). Moreover, the percentage of active β1 integrin nanoclusters was larger than for active β3, in particular inside FAs (30% and 20% for active β1 and active β3, respectively) (Fig. 2A). To further enquire whether both types of integrin nanoclusters act in concert by associating at the nanoscale and sharing their vinculin and talin partners, we performed dual color STORM imaging of the total populations of α5β1 and αvβ3 integrins (Fig. 2C). For this, we used a rabbit anti-α5 antibody (see SI, Table S2) to enable simultaneous labeling with the mouse anti-αvβ3 antibody that we used throughout this study. Visual inspection as well as quantification of the images show that α5β1 and αvβ3 do not intermix at the nanoscale, forming distinct nanoclusters that spatially segregate from each other inside FAs (Fig. S3A, B). The differences in expression levels, activation states, and lateral nano- and meso-scale segregation of α5β1 and αvβ3 suggest the existence of a mechanism that defines the distinct functional roles of these two integrins inside FAs.

To get a better understanding on the global organization of both types of integrins in relation to their main adaptor proteins, we calculated the edge-to-edge distance between the nearest-neighbor integrin nanoclusters and each of the adaptors (see Materials & Methods). Although visual inspection of the dual color STORM images already suggested that integrin nanoclusters are also physically segregated from their adaptor partners (Fig. 1B, C) we used this quantitative approach to estimate the degree of colocalization (i.e., spatial overlap) between different nanoclusters. Since we measure edge-to-edge distances between nearest nanoclusters (EED-NN), values below zero correspond to physically overlapping nanoclusters, EED-NN values above zero to spatially segregated nanoclusters, and EED-NN = 0 corresponds to contiguous nanoclusters, i.e., next to each other. Our results quantitively confirm that the large majority of α5β1 and αvβ3 nanoclusters are laterally segregated from their adaptor partners (Fig. 2D, Fig. S3C, D). Their spatial segregation is essentially determined by the density of nanoclusters within the FAs, as similar EED-NN distributions were obtained from simulations of particles uniformly distributed inside the experimentally obtained FAs, using identical particle densities as to the ones obtained experimentally for the integrin nanoclusters and the different adaptors (Materials & Methods) (Fig. 2D, black lines). Considering the canonical view that integrins inside FAs are bound to paxillin and engaged to talin and vinculin (43, 45), these results might be surprising at first sight, as one would expect a much higher fraction of EED-NN values below zero. Nevertheless, using different super-resolution microscopy techniques, others have also found negligible spatial colocalization of integrins with these adaptor proteins (36, 37). Moreover, Dunn and co-workers also found that most ligand-bound integrins inside FAs exist in a minimally tensioned state (< 2 pN) and not dependent on the actin cytoskeleton (39). The results from the Dunn lab are fully in line with the lack of colocalization of integrins to talin and vinculin as we determined here, since these two adaptors mediate the linkage of integrins to the actin cytoskeleton (43, 46). Nevertheless, we would like to mention that we labeled talin at its actin-binding C-terminal region, which in a 2D projection can appear to be at a distance of 100 nm or more from the N-terminal integrin-binding site (47), so lack of spatial overlap does not necessarily indicate lack of direct engagement.

It has been recently proposed that adhesion proteins including integrins organize in a hierarchical fashion inside FAs, forming nanoclusters similar to the ones reported here, that further enrich into compartmentalized regions around 300 nm in size (37). To enquire whether such hierarchical organization is present in our data, we performed nearest-neighbor distance (NND) analysis between nanoclusters of the same proteins (Fig. 2E). In addition, we implemented simulations of uniformly distributed particles inside the experimental FA regions, using the same experimental nanoclusters densities, and made the corresponding NND analysis (Fig. 2E). Notably, our results show that the global organization of all the adhesion protein nanoclusters investigated is highly uniform inside FAs, as the experimental NND distributions were indistinguishable from those of randomly distributed particles (Fig. 2E). In contrast to the recent findings of the Kusumi lab (37), we did not observe any evidence of spatial enrichment of protein nanoclusters inside FAs. Instead, the observed characteristic separation between nanoclusters appears to be solely determined by their density inside FAs.

Integrin α5β1 but not αvβ3 nanoclusters are enriched on FA edges

Our previous global analysis rendered a uniform inter-cluster organization of all adhesion protein nanoclusters inside FAs. Yet, visual inspection of the dual-color STORM images suggests that a significant fraction of α5β1 nanoclusters preferentially localize on FA edges (Fig. 1B). To quantify these observations, we calculated the shortest-path distance from each α5β1 nanocluster to the edge of its corresponding FA, i.e., its distance-to-FA-edge, and generated a histogram of these values (Fig. 3A). In addition, we generated a similar histogram from simulations of uniformly distributed α5β1 nanoclusters on FAs (see Materials & Methods). In general, if the nanoclusters were uniformly distributed throughout the FA, one would expect that the probability of finding a nanocluster at a specific distance to the FA edge decreases with increasing distance. This occurs because there is a larger area along the perimeter of any given shape and thus a higher number of nanoclusters (e.g., for a circular shape, the decrease is linear with the distance to the FA edge). We found that our experimental and simulated data indeed exhibit this general trend, although, notably, a higher percentage of α5β1 nanoclusters were detected at distances below 100 nm from the FA edges, as compared to a uniform distribution (Fig. 3A and inset), indicative of their preference to locate closer to the FA edges. Similar trends were also obtained for talin, and to a lesser extent for paxillin, whereas the distributions of αvβ3 and vinculin were indistinguishable from random (Fig. S4A).

Integrin α5β1 nanoclusters are preferentially distributed on the edges of FAs.

(A) Distribution of distances from the center of mass (CoM) position of each individual α5β1 nanocluster to its closest FA edge obtained from DBSCAN-analyzed STORM images. The histogram represents the frequency distribution over all cells. Green lines correspond to the experimental data (Exp) and the simulated uniform distribution (Sim) is shown as a black histogram. Bin width is 10 nm. The inset shows the distribution for the first 250 nm. (B) Edge Proximity Factor (EPF) per cell for each of the proteins investigated. Data are shown as median values per individual cell (color dots). The mean ± st. dev. over all the cells are superimposed as black lines. Statistical significance between experimental and simulated data was tested using a paired t-test, where ns = not significant, p>0.05; *, p<0.05; **, p<0.01; ***, p<0.001. (C, D) Variation of the nanocluster distances from the FA edges with respect to random organization, as function of the distance to the FA edges for α5β1 and αvβ3 integrin nanoclusters (C) and for the adaptor proteins (D). (E) Percentage of nanoclusters proximal to FA edges (<100nm) for the total and active β1 nanocluster population (color dots) together with results from simulations of random organization (black dots). Each dot corresponds to the median value per cell, each cell containing multiple FAs. The mean ± st. dev. over all the cells is superimposed as red lines. Statistical significance between experimental and simulated data was tested using a paired t-test, and to test the significance between α5β1 and actβ1 the unpaired t-test was used, assuming a Gaussian distribution of data, where ns = not significant, p>0.05; *, p<0.05; **, p<0.01; ***, p<0.001.

To better assess the extent of nanocluster enrichment at FA edges at the single cell level, we computed the median values of the experimentally obtained distance-to-FA-edge distributions per cell (over multiple FAs) and those corresponding to the simulated data. We then defined the edge-proximity-factor (EPF) as the relative difference between experimental and simulated distributions, where negative EPF values correspond to preferential nanocluster enrichment at FA edges, positive EPF values to nanocluster depletion at FA edges, and EPF = 0 corresponds to random location of nanoclusters with respect to the FA edges (Fig. 3B). Moreover, to statistically determine whether the distribution of EPF values was different from zero (the EPF value expected from a random distance-to-the-FA-edge distribution) we ran a paired Student’s t-test for each protein comparing the experimental with the corresponding simulated distributions. This analysis showed that most proteins exhibit some degree of enrichment at the edges of the FAs (i.e., negative EPF values), although their magnitude was markedly different amongst them (Fig. 3B). The highest enrichments at the FA edges were found for the full population of α5β1 nanoclusters and the active β1 sub-population (∼30% closer to the FA edge than if uniformly distributed) and for talin, whereas the enrichment at the FA edges of the total αvβ3 population and paxillin was much more modest. Notably, the active β3 sub-population and vinculin showed no preferential enrichment at the FA edges (Fig. 3B).

To establish the characteristic distances at which nanocluster enrichment occurs with respect to the FA edges, we next computed the difference between the experimental and simulated distributions, i.e., the variation from random, as a function of the distance to the FA edges, for each of the examined proteins (Fig. 3C, D). Here, positive values correspond to local nanocluster enrichment at a given distance to the FA edge, whereas negative values indicate nanocluster depletion. Our results show a clear difference between both integrins, with α5β1 nanoclusters being enriched close to the edge of FAs (Fig. 3C, continuous green line) whereas αvβ3 nanoclusters showed a rather uniform distribution, regardless of their distance to the FA edges (Fig. 3C, continuous blue line). Moreover, the preferential enrichment of β1 but not of β3 nanoclusters at the FA edges was retained when specifically analyzing the active integrin conformations (Fig. 3C, dashed green and blue lines, respectively).

We next investigated how adaptor protein nanoclusters distribute with respect to the FA edge using a similar approach. We found that both talin and paxillin nanoclusters showed enrichment in a region between 50–150 nm from the FA edges, although somewhat less pronounced as compared to α5β1 nanoclusters (Fig. 3D), whereas vinculin nanoclusters showed no preference to reside proximal to the FA edges. Finally, we extracted the percentage of the total α5β1 and active β1 nanoclusters that reside within distances shorter than 100 nm from the FA edges and compared the results to those from simulations of random organization (Fig. 3E). Notably, a significantly larger percentage of active β1 nanoclusters as compared to the total α5β1 nanocluster population and to the simulations, locate at the periphery of FAs (Fig. 3E).

Together, these results strongly indicate that integrin nanoclusters spatially organize in a rather distinct fashion within FAs, with α5β1 nanoclusters localizing preferentially at the edges of the FAs and αvβ3 nanoclusters distributing randomly inside FAs. Moreover, the preferential enrichment of active β1 nanoclusters together with talin at the FA edges suggests that β1 activation is spatially regulated inside FAs.

Active β1 nanoclusters form multiprotein hubs at FA edges

Since activation of integrins involves engagement with their adaptors paxillin, talin and vinculin, we sought to investigate whether the active β1 nanoclusters preferentially located at the FAs periphery also reside in proximity to their adaptor proteins. For this we first obtained the inter-nanocluster NND (iNND) distributions between active β1 nanoclusters and each of the adaptor nanoclusters (Fig. S4B). We calculated the median of each iNND distribution and used this value as a threshold to classify pairs of active β1-adaptor nanoclusters as being close (i.e., iNND < median) or far (iNND > median) from each other. Next, we separately computed the distance-to-the-FA edge for the active β1 nanoclusters that were found close or far from their partner and compared the results to simulated data (Fig. 4A and inset). Having these plots, we then calculated the variation from random (i.e., difference between experimental data and simulations of random organization) as a function of the distance to the FA edge, for both active β1 nanoclusters close to their partner and active β1 nanoclusters far from their partner (Fig. 4B and Fig. S4C). This analysis shows that active β1 nanoclusters preferentially locate the FA edges regardless of their proximity to talin nanoclusters (Fig. 4B) and to other adaptors (Fig. S4C). In contrast, the same analysis on active β3 nanoclusters shows a random distribution inside of FAs irrespective of their proximity to talin or other adaptor nanoclusters (Fig. 4C).

Active β1 nanoclusters form multiprotein hubs at the FA edges.

(A) Population of active β1-talin nanocluster pairs as a function of their distance-to-the-FA edge. The data plotted in light green correspond to active β1 nanoclusters located close to talin (i.e., iNND < median), while the inset dark green corresponds to active β1 nanoclusters located far from talin (i.e., iNND > median). iNND distributions are shown in Fig. S4B. Black lines are for simulated data sets. Bin width is 20 nm. (B) Variation from random for active β1-talin nanocluster pairs as function of their distance-to-the-FA-edge. Filled bars correspond to active β1 nanoclusters located close to talin (iNND < median) and empty bars correspond to active β1 nanoclusters located far from talin (iNND > median) (C) Similar to (B) but for active β3 nanoclusters. (D-F) Variation from random for paxillin (D), talin (E), vinculin (F) active β1 nanocluster pairs, as function of their distance-to-the-FA-edge, with data segregated in two different distributions according to the pairs proximity. Bin width is 20 nm.

Finally, we enquired whether adaptor protein nanoclusters (paxillin, talin, and vinculin) close to active β1 nanoclusters also locate closer to the FA edges. Notably, we found that all of these adaptor nanoclusters close to active β1 nanoclusters are also enriched on the FA edge region (∼150–200 nm from the edge), whereas adaptor nanoclusters far from active β1 appear to be more uniformly distributed (Fig. 4D–F). Taken together, our data reveal a differential mesoscale organization between active β1 and β3 nanoclusters inside FAs, with active β1 nanoclusters being enriched at the periphery of FAs alongside adhesion adaptor nanoclusters, while active β3 nanoclusters remain uniformly distributed throughout FAs, lacking peripheral enrichment.

Discussion

Here, we have combined dual-color super-resolution optical microscopy together with a robust set of analytical tools to provide insights on the lateral spatial organization of different adhesion proteins inside FAs. Earlier reports showed that nascent adhesions (NAs) are formed due to integrin nanoclusters that either rapidly disassemble at the plasma membrane without force or further mature into FAs in response to mechanical forces (35, 48). Our results now show that even in the presence of mechanical forces (i.e., inside FAs), integrin nanoclusters remain segregated from each other while maintaining their sizes and number of molecules. Similar integrin nanocluster properties were also found outside adhesions, reinforcing the notion that integrin nanoclustering is force-independent (35). Interestingly, we found that the main cytoplasmic adaptors, paxillin, talin and vinculin, also form nanoclusters of similar size and number of single molecule localizations as the integrins, regardless of whether they locate inside or outside FAs. The presence of adhesion protein nanoclusters outside of FAs is intriguing. However, immobilized (active) integrins have been previously observed outside of FAs (25, 49). We suggest that these isolated nanoclusters may represent transient nascent adhesions forming on the cell membrane. Although STORM images do not allow to directly extract the actual number of proteins contained in nanoclusters given the primary and secondary antibody labeling strategy used, the similarity in terms of number of single molecule localizations amongst integrins and their adaptor partners suggest that these nanoclusters might be stoichiometrically similar. Together, our results strongly indicate that adhesion proteins organize as universal nano-hubs, functioning as modular adhesion units to serve multiple integrin-dependent functions related to adhesion and migration but also to mechanosensing and mechanotransduction.

Consistent with previous results (36), our data revealed that active and inactive α5β1 nanoclusters coexist inside FAs but are fully segregated from each other. Remarkably, αvβ3 also exhibits similar nanoscale segregation, with total population-to-active integrin ratios indicating a similar segregation between active and inactive β3. Furthermore, earlier data from our group demonstrated that the integrin αLβ2 expressed on leukocytes also segregates into active and inactive nanoclusters (50, 51). Thus, lateral segregation between active and inactive integrin nanoclusters may represent a common feature to spatially coordinate integrin activation. How could such cooperativity in integrin nanocluster activation arise? Our data show a nearly 1:1 stoichiometry in the number of active integrin nanoclusters, and the main cytoplasmic adaptors talin and vinculin, known to activate integrins by their engagement to the actin cytoskeleton. The availability of a similar number of adaptor nanoclusters containing comparable number of molecules as the integrin nanoclusters suggests that inactive integrin nanoclusters are poised for their collective nanocluster activation, which is achieved upon the concurrent engagement of talin and vinculin nanoclusters (and most probably also paxillin). This mechanism has been shown to operate in NAs in which concurrent recruitment of talin and vinculin prior to force generation is required for integrin activation and further maturation of NAs (52). In addition, our data reveals that there is a large percentage of inactive α5β1 and αvβ3 integrins inside FAs and very little spatial overlap between integrins and main adaptors. As integrin activation is highly dynamic and transient, we propose that this pool of inactive integrin nanoclusters is required to increase the probability of new engagements between integrins and their adaptors once the active integrin nanoclusters become disengaged, stabilizing as a whole the FAs for lifetimes much longer than the binding-unbinding dwell-times of integrins with their partners.

Quantification of super-resolution data showed a higher number of α5β1 nanoclusters compared to their αvβ3 counterparts inside FAs. The relative percentage of active β1 nanoclusters is also higher than that of αvβ3 and moreover these two integrin nanoclusters never intermix inside FAs. Importantly, active β1 nanoclusters preferentially locate at the periphery of FAs and in proximity with paxillin, talin and vinculin nanoclusters, while active β3 are more uniformly distributed inside FAs. The differential degree of α5β1 and αvβ3 in terms of expression, activation state, and lateral nano- and meso-scale segregation, strongly indicates that integrin activation and their distinct functions inside FAs are spatially regulated (Fig. 5). Based on our results we postulate that (i) α5β1, through the ring on the FA periphery and its close proximity to paxillin, talin and vinculin nanoclusters, primarily functions for substrate attachment in its high tensional state, engaging with actomyosin filaments; (ii) αvβ3 and α5β1 integrin nanoclusters in the central region of the FA undertake the role of mechanotransduction, being more dynamic in terms of activation and inactivation states and most probably engaged with shorter branched actin filaments; and (iii) the difference in these roles is controlled by the distinct mechanical forces exerted on the integrins, depending on their location inside FAs, and for α5β1 perhaps also on its different tensional states (53).

Spatial regulation and activation of integrin nanoclusters and protein adaptors inside focal adhesions.

α5β1, αvβ3, and main protein adaptors organize as small nanoclusters inside FAs physically segregated from each other. A considerable high fraction of α5β1 and αvβ3 integrins remain in an inactive state and do not intermix at the nanoscale. Active β1 nanoclusters (green) on the FA edge reside in close proximity to the adaptor protein nanoclusters and thus possibly engaged to actomyosin filaments (red and blue). Integrin nanoclusters in the FA center are more randomly distributed with respect to one another, most probably engaged with branched cortical actin. According to our results we propose that α5β1 nanoclusters located at the FA periphery could be primarily involved in cell substrate attachment, whereas α5β1 and αvβ3 nanoclusters at the center of FAs engage in mechanosensing and mechanotransduction. For simplification, we show a cross-section of a FA where active β1 integrins at the edge are engaged with talin. However, our data also show the presence of active β1 nanoclusters spatially separated from talin and thus possibly engaged to other binding partners linking β1 to the actin cytoskeleton (not shown in the cartoon for the sake of clarity).

Intriguingly, our analysis showed that active β1 nanoclusters preferentially locate the FA edges regardless of their proximity to talin nanoclusters (Fig. 4B). These data are striking since active integrins are expected to be bound to both its ligand and talin, and therefore had anticipated the whole active β1 population to be in close proximity to talin nanoclusters. We can envisage two plausible explanations for active β1 to be far from talin. One possible scenario is the existence of other adaptor protein(s), aside from talin, able to bind to active β1. Such a candidate might be tensin3, which has been recently shown to compete with talin to bind α5β1, during the formation of fAs from FAs (54) A second possibility to explain our observations is the nature of talin itself. As a long rod-like protein, talin stretching experiments have shown that the N- and C-termini can be separated up to 250 nm due to actomyosin activity (55). Since we detect talin by labeling it in its C-terminal domain, which is opposite to the integrin binding side, it is possible that active β1 is engaged to extended talin even for inter-cluster distances larger than the median iNND. Further experiments using specific antibodies against tensin3 as well as labeling talin at the closest proximity to the integrin binding side would clarify the adaptor partners associated to active β1 nanoclusters at the periphery of FAs.

In summary, in contrast to the classical textbook view where FAs are pictured as dense and homogenous assemblies of integrins fully engaged to their cytoplasmic adaptors, our work reveals a highly complex organization inside FAs, where regulation of different integrins and their activation state are highly coordinated in space. We believe that such spatial regulation provides the means for different integrins to synergize and/or diversify their functions in terms of adhesion, migration, mechanosensing and mechanotransduction. Earlier reports showed that proteins such as KANK1, KANK2, and β1 integrins are enriched at the edges of adhesions, with KANK2 reducing the force exerted by the cells through its binding to talin (56, 57). KANK proteins have also been implicated in enabling communication between microtubules and actin by binding to talin and liprin-β1 at FA edges (58). More recent data suggest that the localization of KANK1 around FAs is driven by liquid-liquid phase separation (59). Our findings together with those mentioned above suggest that the α5β1 nanocluster belt is an important feature of FAs and may be critical for their function. However, further research is needed to fully understand how the organization of these key proteins contributes to their roles FAs. p130Cas- and FAK-dependent phase separation has been also suggested to promote integrin clustering and NA assembly (60). Whether phase separation also contributes to the preferential organization and activation of β1 at the FA edges as we reveal here is an intriguing topic that deserves further investigation.

Materials and methods

Sample preparation and immunolabeling

HFF-1 cells (human foreskin fibroblasts) were obtained from ATCC and used with low passage number (passage <15). Cells were maintained in DMEM supplemented with 15% FBS and cultured in a humidified atmosphere at 37° C and 5% CO2. Experiments were performed in cells seeded on 8-well plates (#1 Lab-Tek, Thermofisher, #155411) that were previously coated with fibronectin (FN) at full substrate coverage. FN coating was performed by incubation with 10 µg/ml FN for 30 min at room temperature (RT). We note that FBS can contain vitronectin (VN) and other proteins that may act as ligands of integrins. However, since the glass slides are saturated with FN prior to incubation with cells and media, FN remains the dominant ligand in our experiments. After seeding, cells were allowed to spread for 24 h at 37° C and 5% CO2, after which they were fixed by using 4% PFA in PBS for 15 min at RT. Cells were permeabilized using 0.1% Triton X-100 in PBS for 10 min and blocked with 3% BSA in PBS for 40 min. Primary antibodies against integrins α5β1 and αvβ3, total population, or specific for the active conformation of integrins β1 and β3, and total populations of adaptor proteins, paxillin, talin (actin binding C-terminal epitope), or vinculin, were diluted in blocking buffer and added to the cells for 30 min at RT. After washing with PBS, secondary antibodies that were conjugated in-house with dye pairs Alexa Fluor 405-Alexa Fluor 647 (405-647 channel) or Cy3B-Alexa Fluor 647 (Cy3-647 channel), were added to the cells for 30 min at RT (see Tables S1-S3 for details on reagents and antibodies used). After washing with PBS, cells were stored in PBS until imaged.

Image acquisition

STORM data were obtained on a commercial Nikon Eclipse Ti system (N-STORM), equipped with a 100x oil objective with NA 1.49 using TIRF illumination. The detector was an ANDOR technology EMCCD iXon 897 camera, with a 256×256 pixel ROI and a pixel size of 160 nm. The system has an Agilent technologies laser box with wavelengths of 405 nm, 560 nm and 647 nm. STORM was performed by incubating cells in an imaging buffer containing 100 mM of the reducing agent cysteamine mercaptoethylamine (MEA) (77 mg of MEA in 1 ml of 360 mM HCl, stock solution 1 M), the enzymatic oxygen scavenger Glox solution (0.5 mg/ml glucose oxidase, 40 μg/ml catalase), and 6.25% glucose in PBS. Details of the STORM imaging buffer reagents are found in Table S4. Dual color imaging was carried out cyclically and sequentially. In brief, we first detected the 405-647 channel in a 5-frame cycle consisting of 1 activator frame (405 nm laser on) followed by 4 reporter frames (647 nm laser on), followed by the Cy3-647 channel detection sequence, similarly composed of 1 activator frame (560 nm laser on) followed by 4 reporter frames (647 nm laser on). The frame rate was 50 Hz (20 ms per frame) for all frames. Imaging was carried out by cyclically repeating this 10-frame sequence until all the reporter dyes were exhausted (typically ∼70,000 frames).

Single molecule detection, localization precision, and STORM image reconstruction

Dual-color STORM image reconstruction was done using Insight3 software, provided by Bo Huang (UCSF, initially developed in Xiaowei Zhuang’s lab, Harvard University). In each imaging frame, individual molecules were identified using intensity thresholding (gray levels set between 600-60,000), then each intensity profile was fitted with a 2D Gaussian function to identify the central position, i.e., the coordinates that represent the spatial localization of that fluorescent molecule. The analysis was run over the entirety of each data set, resulting in a final output that is a list of x,y coordinates (localizations) corresponding to each individual molecule that has been identified. Insight3 software allows for drift correction using the correlated function of the image, for details see (61). In this multi-color imaging scheme, the main source of crosstalk is the nonspecific activation of the Alexa 647 by the 647 nm laser (62), which is activator-independent and can occur in any frame. We implemented statistical crosstalk correction where the Insight3 software analyses the full image in 50×50 nm regions of interest (ROIs) and determines if each ROI is predominantly one color of the other. Based on the acquisition design (in our case, one activator frame followed by four reporter frames) the software determines the probability of detecting a nonspecific reporter molecule (62). To minimize color cross-talk, only those localizations detected in the second frame of each five-frame STORM imaging cycle (that is, the first of the four reporter frames) were considered for future analytical steps. Since each single molecule can experience numerous blinking events, the variation of the x,y coordinates from frame to frame determines the effective resolution (63). To measure this, we detected multiple localization events arising from individual molecules deposited on glass and calculated the standard deviation of the localization positions. We measured the effective resolution of our system to be ∼15 nm. These localizations detected from nonspecific antibodies on glass were also used to set a threshold for what we considered to be a nanocluster of proteins. We found that the antibodies on glass had a median number of localizations of 5 (for channel 1) and 3 (for channel 2). With the aim of discarding all localizations coming from nonspecific binding or binding to individual proteins we set the threshold at 10 localizations per nanocluster.

Generation of masks and classification of the localizations belonging to FAs, fAs or outside adhesions

A Fiji plugin, developed by the Lakadamyali lab, was implemented as an initial step to split the localization data from the STORM images into separate data sets corresponding to the localizations contained in certain ROIs (64). First, the data from the two channels of the reconstructed STORM image were imported and the localizations rendered together in a 1024×1024 pixels image (Fig. S1A). Next, we applied an intensity threshold, until the adhesion areas were visually detected, a Gaussian blur of the image (by 2 pixels), and a second threshold filter until a smooth binary mask was obtained (Fig. S1B). The final mask retaining only the adhesion areas was obtained by manually selecting the regions corresponding to the adhesion complexes (identified by visual inspection based on their size and location). In the first steps, this Fiji plugin creates separate files containing the localizations that fall within the selected ROIs: all adhesions, glass, and outside adhesions (i.e., cell membrane that contain the remaining areas of the image that do not correspond to all adhesions or glass). In a second step the “all adhesions” file was again analyzed by the same plugin and all adhesions were split into FAs and fAs and rendered with Insight3. Fig. S1C-F shows the resulting rendered localizations for the different regions. These final two-color files for different cell regions were then individually and independently analyzed.

Image analysis

Clustering detection using the DBSCAN algorithm

DBSCAN is a density-based data clustering algorithm designed to be executed on arbitrary data sets with minimal input from the user (40). The algorithm works by identifying arbitrary-shaped clusters of data points using two user-defined input values: epsilon (ε), which is a radial distance that serves to define a local neighborhood from a randomly selected initial data point; and Nmin, which is the minimum number of points required to be within the initial point’s neighborhood, so it is considered part of the cluster seed. For our image analysis, spots on glass outside the cell regions were analyzed using ε = 20 nm and Nmin = 2, in order to define a minimum threshold for the number of localizations per nanocluster. Important to note that we carried out this analysis on glass areas, next to the cells, found in all of our STORM images. In this way we ensure having the same imaging conditions for in-cell nanoclusters and for spots-on-glass, however, at the expense of having a broader distribution of localization per spot-on-glass due to the presence of cell debris on the glass. When analyzing the nanoclusters found in ROIs within the cell, the DBSCAN parameters were set to ε = 20 nm and Nmin=3 localizations, which are common values used in the field (36, 41). Nevertheless, since multiple blinks of a fluorophore might result in artifactual clusters, we only considered true nanoclusters as those containing 10 or more localizations. Examples on the identification of different nanoclusters by DBSCAN are shown in Fig. S1G. The algorithm also provides the center of mass (CoM) coordinates of each nanocluster together with the number of localizations per nanocluster and their diameter.

Nearest neighbor distance (NND) analysis.

To establish how the protein nanoclusters are distributed relative to one another, we used the MATLAB (MathWorks) function knnsearch. This function computes for each nanocluster the NND, which is the shortest distance from the CoM of a given nanocluster to its closest CoM nanocluster neighbor. We also used the same function to determine the iNND (i.e., NND between nanoclusters of different proteins). To measure the percentage of nanoclusters that overlap (colocalization fraction), we first computed the edge-to-edge distance between nearest neighbors (EED-NN) by subtracting from the NND the radii of the two-corresponding near-neighbor nanoclusters. From the EED-NN distributions we calculated the percentage of occurrences with EED-NN values below zero. The knnsearch function was also used to extract the distance-to-the-FA edge, i.e., the shortest distance between nanoclusters and the edge of adhesions. To that end, we computed the NND between the CoM of each nanocluster to the center coordinates of the list of pixels that fall on the edge of the FA mask.

Computational generation of uniform nanocluster distributions

To assess whether the distance distributions obtained from the experimental data (NND, EED-NN and distance-to-the-FA-edge), correspond to a preferential type of organization or to a uniform distribution, we performed simulations by generating in silico histograms of non-overlapping and randomly distributed nanoclusters inside FAs. For this, we first extracted the exact number and size of nanoclusters per adhesion from the experimental data. Although our initial approach was to computationally generate the same number of nanoclusters and distribute them uniformly over the adhesion masks regardless of their physical size, we found that this approach was not computationally suitable. The reason for it is that in many cases the large nanoclusters did not fit (without overlapping) in the remaining space available within the adhesions after randomly positioning the large majority of smaller nanoclusters. We thus opted for classifying the nanoclusters according to their physical size and then start our procedure by first positioning the largest nanoclusters and later all the remaining smaller ones. In addition, as our experimental data renders well segregated nanoclusters, we imposed in our algorithm the condition that nanoclusters should not overlap with one another. Fig. S5 shows the different steps followed in order to generate distributions of random nanoclusters inside FAs.

Statistical analysis

Statistical analysis was carried out using GraphPad Prism (version 9.12). The details of the individual statistical tests as well as the level of significance along with p-values are explicitly provided in the figure legends. No predetermined statistical method was employed for establishing the sample size, as the sample sizes were not selected based on a predefined effect size. In this context, multiple experiments were independently performed, employing various sample replicates, as detailed in Table S5.

eLife Assessment

In this manuscript, the authors analyse the nanoscale localisation of α5β1 and αVβ3 integrins in integrin adhesion complexes (IAC) by dual-colour STORM and assess the spatial organisation at the nano and mesoscale of their main adaptors (paxillin, talin and vinculin). This is an important work that provides detailed analyses that reveal how elements of these complex structures are really organised at the nanoscale, an essential perspective for a better understanding of how IACs function and regulate mechanotransduction processes. The evidence presented is solid, with super-resolution imaging experiments conducted using a single, validated methodology and subsequent computational modelling that enabled a quantitative assessment of the resulting data.

Reviewer #1 (Public review):

Summary:

In recent years, it has become increasingly evident how beautifully intricate IAC are at the nanoscale. Studies like the one presented here that shed light on the precise inner organisation of IAC are thus quite important and relevant in order to obtain a better in-depth understanding of IAC functioning and the contribution of different integrin subtypes to cell adhesive and mechanotransductive processes.

Interestingly, the authors found a distinct localisation of α5β1 and αVβ3 integrin nanoclusters within focal adhesion of human fibroblasts, with α5β1 integrin nanoclusters being at the periphery of IAC and αVβ3 integrin nanoclusters randomly distributed. Furthermore, a surprisingly high percentage of inactive integrins within IAC and relatively low spatial integrin colocalisation with adaptor proteins has been shown.

Strengths:

This is a very thoroughly performed STORM-based assessment of the nanodistribution of α5β1 and αVβ3 nanoclusters within IAC (and outside). The image quality is outstanding, and the authors have meticulously executed the experiments and the image analyses.

Weaknesses:

The only weakness is maybe that the manuscript remains descriptive. However, the high quality of the "description" of the nano-organisation of IAC by this scrupulous study is really important to better understand the inner workings of IAC. It provides a very solid foundation to look deeper into the (patho)physiological implications of this organisation, see recommendations (which are rather suggestions in this case).

Reviewer #2 (Public review):

Summary:

In this study, dual-color super-resolution microscopy analysis was performed to study the co-operation between integrins and focal adhesion proteins in human fibroblast cells. The study focused on two integrins which have been previously found to be mainly responsible for focal adhesions, namely α5β1 and αvβ3.

Specifically, the study tried to shed light on the nanoclustering of integrins in focal adhesions.

In the current study, more integrin nanoclusters were observed in focal adhesions compared to other cell-matrix adhesion structures. The study revealed that both α5β1 and αvβ3 form nanoclusters, and those appear segregated from each other. While αvβ3 nanoclusters organize randomly inside focal adhesions regardless of their activation state, α5β1 nanoclusters, and particularly the nanoclusters containing β1-integrin in active conformation, preferentially organized at the edges of focal adhesions. The nanoclusters formed by each integrin were similar in size.

Cytoplasmic adapter proteins appeared less in nanocluster assemblies, suggesting that integrin nanoclusters are also forming without the studied cytoplasmic adapter proteins (talin, vinculin, paxillin). Active integrins were identified with the help of conformation-specific antibodies, and this enabled us to study the colocalization between integrins and their cytoplasmic adapter proteins. This analysis revealed that activated integrins are strongly engaged with adapter proteins

Strengths:

The study stems from the thorough computational modelling of the nanoclusters, which enables quantification of the behavior of the clusters, including their mesoscale distribution.

The study strengthens the view that α5β1 and αvβ3 have specific functions in focal adhesions, α5β1 nanoclusters localizing preferentially on focal adhesion edges. The study also revealed that nanoclusters localized at the edges of focal adhesion were enriched for talin and paxillin but not for vinculin.

Analysis of adaptor protein nanoclusters (paxillin, talin, and vinculin) revealed that all adapter protein nanoclusters studied here close to active β1 nanoclusters are enriched on the focal adhesion edge region, whereas integrin adaptor nanoclusters far from active β1 appear to be more uniformly distributed.

Importantly, the current study suggests that integrin subtype-specific nanoclusters are not only present at an early stage of adhesion formation, but integrin nanoclusters remain segregated from each other also in mature focal adhesions, maintaining their sizes and number of molecules.

Interestingly, the study revealed that selected cytoplasmic adaptors (paxillin, talin, and vinculin), also form nanoclusters of similar size and number of single molecule localizations as the integrins, regardless of whether they locate inside or outside focal adhesions. The adapter nanoclusters are enriched in the focal adhesion "belt", colocalizing with the active α5β1 integrin nanoclusters.

Weaknesses:

The current study is highly dependent on the antibodies. It is possible that antibodies containing two binding sites for antigen influence the nanoscale organization (and also activation) of the receptors. Control experiments to study the possible contribution of antibodies to the measured outcome should be performed to verify the main findings. One possible approach could be to use fluorescently tagged integrins available. Alternatively, integrins (or adapter proteins) could be tagged with a small ligand and detected using a monovalent binder.

Only a limited number of integrin adapter proteins were investigated. Given the high number of identified adapter proteins, this is an understandable choice. However, it would be fascinating to understand if the nanoclusters of inactive integrins are dominantly bound with a certain adapter protein, such as tensin.

Reviewer #3 (Public review):

Summary:

In their study, the authors reveal using dual-color super-resolution STORM microscopy modality and immunolabeling in fixed adherent cells, that β1 and β3 integrins as well as adaptors (paxillin, talin and vinculin) are all organized in nanoclusters of similar size (50nm) and molecular density (20 copy number) inside FAs but also outside. Using activity-specific immunolabeling of β1 and β3 integrins, they revealed that active integrin subpopulations were both clustered but in distinct exclusive nano-aggregates in agreement with Spiess et al. (2018). Once more, the "active" integrin nanoclusters displayed similar properties in terms of size and molecular density, suggesting that molecular organization in nanoclusters is an intrinsic property of integrins in plasma membrane multimerizing independently of their location (inside or outside FAs), their level of activation, or their connection to the cytoskeleton. Then the authors followed up by analyzing at the mesoscale how these "universal" nanoclustered adhesive units are distributed spatially. Inspecting the surface density of nanoclusters revealed that the density of integrin nanoclusters in FAs was 5x larger, compared to integrin nanoclusters outside adhesions. Interestingly, whereas the density of total integrin nanoclusters was 2-4x larger than adaptor nanoclusters, the density of "active" integrin nanoclusters stoichiometrically matches that of talin and vinculin nanoclusters, and was slightly outnumbered by paxillin nanoclusters. These findings suggest that inside FAs, among the total number of integrin nanoclusters, the subset of "active" integrin nanoclusters could be engaged with "adaptor" nanoclusters on a 1:1 ratio. Using analysis of the nearest neighbor distance (NND) between distinct integrin clusters and each of the adaptors, the authors report that they found negligible spatial colocalization of integrins with these adaptor proteins and that spatial segregation is essentially determined by the density of nanoclusters within the FAs. As authors reported that α5β1 and αvβ3 do not intermix at the nanoscale, the authors finally highlighted how α5β1 and αvβ3 distinct nanoclusters are differently organized and segregated inside FAs. Adapting the NND analysis in order to inspect how far the nanoclusters are from the edges of FAs they are located in, authors revealed that α5β1 but not αvβ3 integrin nanoclusters are enriched on FA edges and that similar FA edge-enriched distribution for "active" α5β1 and adaptor protein nanoclusters was found for talin and paxillin but not vinculin. The latter results suggest that FA edges could constitute multiprotein hubs for enhanced colocalization and activation for α5β1 integrin nanoclusters and adaptors such as talin and paxillin. Unfortunately NND analysis could not confirm this enhanced colocalization hypothesis.

General Assessment:

While the study presents some valuable findings, it reads currently as a compilation of intriguing but preliminary observations derived primarily from a single methodology (dual-color STORM and DBSCAN clustering analysis). As the initial findings often lack confirmation through additional data analysis (such as the NND analysis the authors used), there's a critical necessity to bolster the methodological approach. This should involve replicating the main findings using alternative single-molecule super-resolution techniques (such as quantitative DNA-PAINT) or employing different clustering analytical tools (such as voronoi-tessellation). Furthermore, the manuscript feels incomplete, focusing solely on describing molecular organization without offering substantial insights into how these observations correlate with the regulation, activation, and functionality of integrins at the cellular level.

The manuscript presents extensive datasets and utilizes methodologies in which the investigators demonstrate expertise. Nevertheless, there's uncertainty regarding the novelty and broad appeal of the findings. For instance, the observation of integrin nanoclustering has been previously reported in several publications (e.g., Changede et al., Dev Cell 2015; Spiess et al., JCB 2018; Fujiwara et al., JCB 2023). Similarly, the accumulation of specific proteins at the periphery of FAs has been documented elsewhere (e.g., Sun et al., NCB 2016; Stubb et al., NatComm 2019; Nunes-Vicente TCB 2023), as well as the differential dynamic organization of α5β1 and αvβ3 integrins inside FAs (e.g., Rossier et al., NCB 2012). Beyond the universal organization of adhesive proteins, there's a need to identify novel insights that significantly advance the field. One potential avenue could involve pinpointing the molecular determinant controlling the FA edge enrichment of active α5β1 integrins and talin nanoclusters. For instance, could there be an interplay between α5β1 and αvβ3 integrin nanoclusters visible on one's organisation when suppressing the other using deletion (KO) or depletion (SiRNA)? Also, could KANK, which also exhibits enrichment and regulates talin activity (e.g., Sun et al., NCB 2016), play a role in this process? Identifying the molecular players that regulate even partially the mesoscale organization of nanoclusters of proteins would really benefit the breadth of this manuscript.

Echoing the previous concern, the manuscript described a novel and rather surprising finding related to molecular clustering of adhesion proteins. Indeed, the fact that nanoclusters exhibit uniform size and molecular density regardless of the protein type, location, or activation level is indeed surprising and raises many questions about the methodology used to assess molecular clustering. I feel that the description and characterization of integrin nanoclusters appear incomplete and need to be expanded by comparing different analytical strategies for protein clustering. Furthermore, a lack of the manuscript in its actual form concerns the quantification of integrin numbers inside the observed nanoclusters. I agree that the path from optical microscopy to protein stoichiometry quantification is hard and full of drawbacks. But the authors do not fully address these issues that are extremely important when discussing protein nanoclustering. This quantitative aspect should be discussed.

First, it is crucial for the authors to carefully examine and discuss in their manuscript whether there are any potential biases or limitations in the experimental techniques (dual-color STORM) or data analysis methods employed (DBSCAN). Second, the authors did not in the current manuscript, but should provide control samples to demonstrate the sensitivity and dynamic range of their experimental strategy.

In STORM images displayed in Figure S1, the authors highlighted localization clusters detected by DBSCAN as a signature for integrin nanoclusters. But the authors do not discuss the localization spots that were not detected by DBSCAN. Could they be individual integrins? And if so, they should also be considered as useful information? This brings me to another related technical question about how DBSCAN handles the case where fluorescent molecules are blinking. This is important as multiple emissions by a single fluorophore could be detected as a nanocluster of several molecules where it would be an artefact due to the photophysics of the fluorophore. Could the authors comment on these points?

Also, using isolated and stochastically physisorbed fluorophores (Ab coupled with activator /reporter pairs used in this study) on glass helped define the signature in STORM of a single isolated molecule. To obtain the signature of clustered fluorophores, the authors could use anti-donkey antibodies to cross-link those STORM-specifically labeled Ab as a means to artificially obtain clustered fluorophores. Ultimately, to avoid the bias effect of the glass surfaces on the photophysics of fluorophores and be in the same imaging conditions as for the described nanoclusters, the authors should use model systems composed of multimers of GFP vs. single GFP, immunolabeled with a GFP-binding monoclonal antibody. This will permit evaluation of the cluster signature obtained with DBSCAN analysis of STORM data for single vs. multimers of known stoichiometry. This would constitute an undisputable molecular stoichiometry ruler.

Due to the surprising finding of the nanoclusters' "universality", it is imperative for the authors to validate the findings through complementary methodologies and analytical tools. This should involve replication of results using alternative super-resolution techniques (quantitative DNA-PAINT) and exploring different clustering algorithms (Voronoï-Tesselation) to ensure the robustness and reliability of the observations.

Author response:

The following is the authors’ response to the original reviews.

We have addressed all the concerns and recommendations by the reviewers, in particular, the requested control experiments using alternative super-resolution microscopy approaches and analysis of the data using Voronoi tessellation in addition to DBSCAN, as requested by reviewer 3. We also provide additional data on tensin3 as suggested by reviewer 2. Finally, to provide a first insight on the role of mechanical forces in the distribution of integrin nanoclusters inside FAs as recommended by reviewer 1, we have performed experiments at different cell seeding times where it is known that FA maturation over time requires mechanical forces.

Public Reviews:

Reviewer #1 (Public review):

Summary:

In recent years, it has become increasingly evident how beautifully intricate IAC are at the nanoscale. Studies like the one presented here that shed light on the precise inner organisation of IAC are thus quite important and relevant in order to obtain a better in-depth understanding of IAC functioning and the contribution of different integrin subtypes to cell adhesive and mechanotransductive processes.

Interestingly, the authors found a distinct localisation of α5β1 and αvβ3 integrin nanoclusters within focal adhesion of human fibroblasts, with α5β1 integrin nanoclusters being at the periphery of IAC and αvβ3 integrin nanoclusters randomly distributed. Furthermore, a surprisingly high percentage of inactive integrins within IAC and relatively low spatial integrin colocalisation with adaptor proteins has been shown.

Strengths:

This is a very thoroughly performed STORM-based assessment of the nanodistribution of α5β1 and αvβ3 nanoclusters within IAC (and outside). The image quality is outstanding, and the authors have meticulously executed the experiments and the image analyses.

We are grateful to the reviewer for acknowledging the strengths of our study.

Weaknesses:

The only weakness is maybe that the manuscript remains descriptive. However, the high quality of the "description" of the nano-organisation of IAC by this scrupulous study is really important to better understand the inner workings of IAC. It provides a very solid foundation to look deeper into the (patho)physiological implications of this organisation, see recommendations (which are rather suggestions in this case).

We thank the reviewer for their feedback and have addressed their recommendations in our updated manuscript and accompanying reply (see recommendations to the authors). In summary, we have now performed experiments at different seeding times as FA maturation requires mechanical forces, and enquired whether forces might play a role in establishing the spatial distribution of the two different integrins within more mature IACs. The results are now shown as new Fig. 2 and discussed in pages 9 and 10. In addition, in order to get a first insight into the biological implications of our findings we performed dual-colour super-resolution experiments of tensin-3 and α5β1 in FAs, as tensin-3 has been implicated in fibronectin fibrillogenesis. The results are now shown in Fig. S8 and we discuss their potential implications in pages 22 and 23 of the revised manuscript (see more details in the reply to the recommendation to the authors).

Reviewer #2 (Public review):

Summary:

In this study, dual-color super-resolution microscopy analysis was performed to study the co-operation between integrins and focal adhesion proteins in human fibroblast cells. The study focused on two integrins which have been previously found to be mainly responsible for focal adhesions, namely α5β1 and αvβ3.

Specifically, the study tried to shed light on the nanoclustering of integrins in focal adhesions.

In the current study, more integrin nanoclusters were observed in focal adhesions compared to other cell-matrix adhesion structures. The study revealed that both α5β1 and αvβ3 form nanoclusters, and those appear segregated from each other. While αvβ3 nanoclusters organize randomly inside focal adhesions regardless of their activation state, α5β1 nanoclusters, and particularly the nanoclusters containing β1-integrin in active conformation, preferentially organized at the edges of focal adhesions. The nanoclusters formed by each integrin were similar in size.

Cytoplasmic adapter proteins appeared less in nanocluster assemblies, suggesting that integrin nanoclusters are also forming without the studied cytoplasmic adapter proteins (talin, vinculin, paxillin). Active integrins were identified with the help of conformation-specific antibodies, and this enabled us to study the colocalization between integrins and their cytoplasmic adapter proteins. This analysis revealed that activated integrins are strongly engaged with adapter proteins.

Strengths:

The study stems from the thorough computational modelling of the nanoclusters, which enables quantification of the behavior of the clusters, including their mesoscale distribution.

The study strengthens the view that α5β1 and αvβ3 have specific functions in focal adhesions, α5β1 nanoclusters localizing preferentially on focal adhesion edges. The study also revealed that nanoclusters localized at the edges of focal adhesion were enriched for talin and paxillin but not for vinculin.

Analysis of adaptor protein nanoclusters (paxillin, talin, and vinculin) revealed that all adapter protein nanoclusters studied here close to active β1 nanoclusters are enriched on the focal adhesion edge region, whereas integrin adaptor nanoclusters far from active β1 appear to be more uniformly distributed.

Importantly, the current study suggests that integrin subtype-specific nanoclusters are not only present at an early stage of adhesion formation, but integrin nanoclusters remain segregated from each other also in mature focal adhesions, maintaining their sizes and number of molecules.

Interestingly, the study revealed that selected cytoplasmic adaptors (paxillin, talin, and vinculin), also form nanoclusters of similar size and number of single molecule localizations as the integrins, regardless of whether they locate inside or outside focal adhesions. The adapter nanoclusters are enriched in the focal adhesion "belt", colocalizing with the active α5β1 integrin nanoclusters.

We are grateful to the reviewer for acknowledging the strengths of our study.

Weaknesses:

The current study is highly dependent on the antibodies. It is possible that antibodies containing two binding sites for antigen influence the nanoscale organization (and also activation) of the receptors. Control experiments to study the possible contribution of antibodies to the measured outcome should be performed to verify the main findings. One possible approach could be to use fluorescently tagged integrins available. Alternatively, integrins (or adapter proteins) could be tagged with a small ligand and detected using a monovalent binder.

We understand the concern of the reviewer regarding the use of antibodies for imaging. Nevertheless, we would like to clarify that antibody labelling has always been performed after cell fixation, precluding potential cross-linking artefacts due to protein mobility and avoiding unwanted receptor activation.

Nevertheless, and although it is highly unlikely to happen in fixed cells, there could be two potential sources of antibody (Ab) labelling artefacts. As the reviewer noted, a primary Ab containing two binding sites could bind to two adjacent proteins (within ~10 nm from each other), potentially underestimating the stoichiometry of the nanoclusters, i.e., number of receptors or proteins per nanocluster. However, in our manuscript we never attempted to provide an estimation of the nanocluster stoichiometry, as it is highly challenging (and prone to artefacts) to provide quantification of the number of proteins using super-resolution-based single-molecule localisation methods which rely on the stochastic blinking of individual fluorophores.

A second source for potential artefacts comes from the use of the secondary Ab, which (albeit unlikely) could bind to two different primary Abs. To exclude this potential artefact, we performed super-resolution imaging using DNA-PAINT as a different imaging strategy. In this case, the DNA docking site is site-specifically coupled to one camelid single-domain Ab (sdAB), having a much smaller size as compared to a secondary Ab, reducing therefore linkage error and increasing the accessibility of primary Ab-labelled proteins. These new data are included now in Fig. S4. As can be observed, no differences in terms of nanocluster sizes and/or compositions were observed for any of the proteins investigated using DNA-PAINT as compared to our initial STORM data. These control experiments thus rule out any potential artefacts introduced by the secondary Ab (for more details, please see the reply to the recommendations for authors section).

Only a limited number of integrin adapter proteins were investigated. Given the high number of identified adapter proteins, this is an understandable choice. However, it would be fascinating to understand if the nanoclusters of inactive integrins are dominantly bound with a certain adapter protein, such as tensin.

We fully agree with the reviewer and have now performed dual-colour super-resolution STED microscopy of α5β1 and tensin-3 on HFF cells seeded for 24 hours. Interestingly, instead of being an integrin inactivator, we found that tensin-3 is also highly enriched at the FA periphery where a large fraction of active β1 integrins are located, suggesting that at these particular regions, active β1 could be either engaged to talin (as shown in our original data) or to tensin-3 (our new data shown in Fig. S8). We provide more details of our answer in the section of “recommendation to the authors”. Additional experiments, which in our opinion fall outside of the scope of this work, would be necessary to identify other potential integrin inactivator partners, but certainly a topic of future interest to our group.

Reviewer #3 (Public review):

Summary:

In their study, the authors reveal using dual-color super-resolution STORM microscopy modality and immunolabeling in fixed adherent cells, that β1 and β3 integrins as well as adaptors (paxillin, talin and vinculin) are all organized in nanoclusters of similar size (50nm) and molecular density (20 copy number) inside FAs but also outside. Using activityspecific immunolabeling of β1 and β3 integrins, they revealed that active integrin subpopulations were both clustered but in distinct exclusive nano-aggregates in agreement with Spiess et al. (2018). Once more, the "active" integrin nanoclusters displayed similar properties in terms of size and molecular density, suggesting that molecular organization in nanoclusters is an intrinsic property of integrins in plasma membrane multimerizing independently of their location (inside or outside FAs), their level of activation, or their connection to the cytoskeleton. Then the authors followed up by analyzing at the mesoscale how these "universal" nanoclustered adhesive units are distributed spatially. Inspecting the surface density of nanoclusters revealed that the density of integrin nanoclusters in FAs was 5x larger, compared to integrin nanoclusters outside adhesions. Interestingly, whereas the density of total integrin nanoclusters was 2-4x larger than adaptor nanoclusters, the density of "active" integrin nanoclusters stoichiometrically matches that of talin and vinculin nanoclusters, and was slightly outnumbered by paxillin nanoclusters. These findings suggest that inside FAs, among the total number of integrin nanoclusters, the subset of "active" integrin nanoclusters could be engaged with "adaptor" nanoclusters on a 1:1 ratio. Using analysis of the nearest neighbor distance (NND) between distinct integrin clusters and each of the adaptors, the authors report that they found negligible spatial colocalization of integrins with these adaptor proteins and that spatial segregation is essentially determined by the density of nanoclusters within the FAs. As authors reported that α5β1 and αvβ3 do not intermix at the nanoscale, the authors finally highlighted how α5β1 and αvβ3 distinct nanoclusters are differently organized and segregated inside FAs. Adapting the NND analysis in order to inspect how far the nanoclusters are from the edges of FAs they are located in, authors revealed that α5β1 but not αvβ3 integrin nanoclusters are enriched on FA edges and that similar FA edge-enriched distribution for "active" α5β1 and adaptor protein nanoclusters was found for talin and paxillin but not vinculin. The latter results suggest that FA edges could constitute multiprotein hubs for enhanced colocalization and activation for α5β1 integrin nanoclusters and adaptors such as talin and paxillin. Unfortunately NND analysis could not confirm this enhanced colocalization hypothesis.

General Assessment:

While the study presents some valuable findings, it reads currently as a compilation of intriguing but preliminary observations derived primarily from a single methodology (dual-color STORM and DBSCAN clustering analysis). As the initial findings often lack confirmation through additional data analysis (such as the NND analysis the authors used), there's a critical necessity to bolster the methodological approach. This should involve replicating the main findings using alternative single-molecule super-resolution techniques (such as quantitative DNA-PAINT) or employing different clustering analytical tools (such as voronoi-tessellation). Furthermore, the manuscript feels incomplete, focusing solely on describing molecular organization without offering substantial insights into how these observations correlate with the regulation, activation, and functionality of integrins at the cellular level.

We appreciate the comment of the reviewer and have taken their recommendation to heart in order to validate our methodology. In summary, we have now performed extensive DNA-PAINT to replicate most of our initial findings obtained by STORM, as requested by the reviewer. In addition, as a different super-resolution imaging strategy, we have also used STED microscopy to confirm the nanoclustering of integrins and some of the adaptors demonstrating now, by means of three different super-resolution techniques, that both integrins and their adaptors form nanoclusters of similar size and composition, regardless of whether they are inside or outside FAs. We have included these data as Figs. S3 and S4 and discussed the results in pages 8-9 of the main manuscript.

Regarding the use of an alternative analysis for the data, we have now used the Voronoi tessellation algorithm to re-analyse our STORM data, as requested by the reviewer. The results of the analysis, which render similar sizes and number of localizations as obtained by DBSCAN, are now included in Fig. S5 and mentioned in page 8 of the main manuscript.

The manuscript presents extensive datasets and utilizes methodologies in which the investigators demonstrate expertise. Nevertheless, there's uncertainty regarding the novelty and broad appeal of the findings. For instance, the observation of integrin nanoclustering has been previously reported in several publications (e.g., Changede et al., Dev Cell 2015; Spiess et al., JCB 2018; Fujiwara et al., JCB 2023). Similarly, the accumulation of specific proteins at the periphery of FAs has been documented elsewhere (e.g., Sun et al., NCB 2016; Stubb et al., NatComm 2019; Nunes-Vicente TCB 2023), as well as the differential dynamic organization of α5β1 and αvβ3 integrins inside FAs (e.g., Rossier et al., NCB 2012). Beyond the universal organization of adhesive proteins, there's a need to identify novel insights that significantly advance the field. One potential avenue could involve pinpointing the molecular determinant controlling the FA edge enrichment of active α5β1 integrins and talin nanoclusters. For instance, could there be an interplay between α5β1 and αvβ3 integrin nanoclusters visible on one's organisation when suppressing the other using deletion (KO) or depletion (SiRNA)? Also, could KANK, which also exhibits enrichment and regulates talin activity (e.g., Sun et al., NCB 2016), play a role in this process? Identifying the molecular players that regulate even partially the mesoscale organization of nanoclusters of proteins would really benefit the breadth of this manuscript.

We could not agree more with the reviewer and in fact, we are currently investigating the mechanisms that control the enrichment of α5β1 and adaptors at the edges of FAs. However, considering the amount of work needed to determine the spatiotemporal organization of other molecular players using super-resolution imaging constitutes a major tour de force.

To get a first insight into the process of active α5β1 enrichment at the FA edges, we hypothesised that mechanical forces exerted by the actomyosin machinery could influence the lateral distribution of both integrin subsets (α5β1 and αvβ3) inside FAs. Since FA maturation and strengthening over time requires mechanical forces, we performed experiments at different cell seeding times (90 min, 3 hours and 24 hours) and used STORM imaging to follow the evolution of integrin nanoclustering in time as well as their spatial distributions inside FAs. Interestingly, while nanoclustering of both integrin sub-sets inside FAs is not influenced by seeding times, their lateral distribution was markedly different, with α5β1 nanocluster distribution being already established at earlier seeding times, while αvβ3 nanocluster distribution appeared as rather random at earlier seeding times and progressively organized reaching a well-defined lateral spacing at 24 hours of spreading time. These initial data strongly suggest that mechanical forces might play a role in the distinct lateral distribution of both subsets of integrin nanoclusters over time. We have now included these data as new Fig. 2 of the revised manuscript and discuss the results in the associated text (pages 9 and 10). We also discuss potential avenues for further research along the directions suggested by the reviewer.

In addition, since it has been recently shown that tensin-3 interaction with talin drives the formation of fibronectin-associated fibrillar adhesions (Atherton et al, J Cell Biol 2022) which are enriched in β1 integrins, we performed dual-colour super-resolution STED microscopy of β1 and tensin-3 on HFF cells seeded for 24 hours. Interestingly, our initial data show co-enrichment of both tensin-3 and active β1 nanoclusters at the FA periphery, suggesting that at these particular regions, active β1 could be either engaged to talin (as shown in our original manuscript) or to tensin. Our current working hypothesis is that α5β1 enrichment at the FA periphery serves to facilitate the translocation of α5β1 integrins from FAs to fibrillar adhesions, most probably in a talin-tensin-dependent manner. We have now included these data as Fig. S8 and accompanying discussion in pages 22 and 23 of the revised manuscript.

Echoing the previous concern, the manuscript described a novel and rather surprising finding related to molecular clustering of adhesion proteins. Indeed, the fact that nanoclusters exhibit uniform size and molecular density regardless of the protein type, location, or activation level is indeed surprising and raises many questions about the methodology used to assess molecular clustering. I feel that the description and characterization of integrin nanoclusters appear incomplete and need to be expanded by comparing different analytical strategies for protein clustering. Furthermore, a lack of the manuscript in its actual form concerns the quantification of integrin numbers inside the observed nanoclusters. I agree that the path from optical microscopy to protein stoichiometry quantification is hard and full of drawbacks. But the authors do not fully address these issues that are extremely important when discussing protein nanoclustering. This quantitative aspect should be discussed.

We appreciate the comment of the reviewer as indeed, the existence of “universal” nanoclusters is intriguing. Recently, together with Prof. S. Mayor we have written a short review in Curr. Opin. Cell Biol 2024 proposing that nanoclustering constitutes a molecular-scale organisation principle that governs cellular information flow at the plasma membrane. Our proposal is supported by an extensive number of recent papers showing that most cell membrane receptors and downstream signalling components are organized as pre-assembled nanoclusters. We posit that these nanoclusters serve as modular units whose concatenation in a specific spatiotemporal sequence leads to distinct signalling outputs. Thus, the existence of universal nanoclusters of integrin receptors and adaptors is indeed intriguing but not surprising to us.

In any case, the concern of the reviewer is well-taken, and as mentioned above, we have used a different algorithm to detect and quantify nanoclustering, obtaining similar values using either Voronoi tessellation or DBSCAN approaches. These data are now included as Fig. S5 in the manuscript.

Regarding the quantification of integrin numbers inside the observed nanoclusters, we agree with the reviewer that determining protein stoichiometry using single-molecule localization microscopy or STED remains a major technical challenge and is highly prone to artefacts. For this reason, we refrain from making claims about absolute protein numbers per nanocluster. Our relative comparison of nanoclustering among the different proteins investigated is thus exclusively based on the number of single-molecule localisations contained in each nanocluster which is a fair approach since we always use the same reporter fluorophore and maintain similar excitation conditions throughout our experiments. We have now included a few lines on page 9 regarding quantification of the absolute protein numbers inside the nanoclusters and further discuss in the revised manuscript the limitations of single-molecule localisation methods towards the stoichiometry determination of the nanoclusters (see page 20 of the revised manuscript).

First, it is crucial for the authors to carefully examine and discuss in their manuscript whether there are any potential biases or limitations in the experimental techniques (dual-color STORM) or data analysis methods employed (DBSCAN). Second, the authors did not in the current manuscript, but should provide control samples to demonstrate the sensitivity and dynamic range of their experimental strategy.

As already mentioned, we have validated the STORM data using both DNA-PAINT and STED and, validated our data analysis obtained with DBSCAN using the Voronoi tessellation algorithm. See Figs. S3, S4 and S5. In terms of sensitivity and dynamic range of our methodology: our set-up has single-molecule detection sensitivity which is demonstrated by the fact that we observe and detect discrete blinking events, a property of single-molecule fluorescence emission and key ingredient to super-resolution single-molecule localisation microscopy. The dynamic range (if we understand correctly the question of the reviewer) is given by the number of frames used to accumulate single-molecule localisations. In our case, we stop acquisition after we deplete most of the single-molecule spots in the imaging view, which typically occurred after 70,000 frames acquisition, as correctly mentioned in the material & methods section.

In STORM images displayed in Figure S1, the authors highlighted localization clusters detected by DBSCAN as a signature for integrin nanoclusters. But the authors do not discuss the localization spots that were not detected by DBSCAN. Could they be individual integrins? And if so, they should also be considered as useful information? This brings me to another related technical question about how DBSCAN handles the case where fluorescent molecules are blinking. This is important as multiple emissions by a single fluorophore could be detected as a nanocluster of several molecules where it would be an artefact due to the photophysics of the fluorophore. Could the authors comment on these points?

As mentioned in the original manuscript, between 20-30% of the localizations were not assigned to nanoclusters (Fig. S1H, I) since we imposed a minimum of ten localizations within the radius defined by DBSCAN to be considered as a true nanocluster. This essentially means that regions with less than 10 localizations were not considered in our nanoclustering analysis. However, we cannot be certain as to whether these lower number of localizations correspond to individual integrins, stochastic blinking of the fluorophore or small aggregates containing only a couple of integrins, for the same reasons that we cannot provide quantification of the absolute number of proteins included in each nanocluster: stoichiometry determination by means of single-molecule super-resolution methods is highly prone to artefacts.

Regarding the concern of how DBSCAN handles fluorophore blinking, the reviewer is completely right as the photophysics of the fluorophore can influence the analysis of the data and the identification of true nanoclusters. To decouple the photophysics of the fluorophore we first assess the number of blinking events within the DBSCAN radius, i.e., number of localizations corresponding to individual antibodies sparsely distributed on the glass surface. In our case, the median values for the two activator-reporter pairs corresponded to 5 localizations for Alexa 405-Alexa647-conjugated Abs and 3 localizations for Cy3-Alexa 647-conjugated Abs (see Fig. 1E). Yet, despite these median values, the number of localizations per individual Ab naturally shows a distribution. Thus, to avoid any overestimation in the degree of nanoclustering, we impose an additional constrain to our analysis and consider true nanoclusters only those ones containing at least 10 localizations. We have now significantly extended the explanation in the main text (see page 6) as well as materials & methods so that it becomes clearer to the reader.

Also, using isolated and stochastically physisorbed fluorophores (Ab coupled with activator /reporter pairs used in this study) on glass helped define the signature in STORM of a single isolated molecule. To obtain the signature of clustered fluorophores, the authors could use anti-donkey antibodies to cross-link those STORM-specifically labeled Ab as a means to artificially obtain clustered fluorophores. Ultimately, to avoid the bias effect of the glass surfaces on the photophysics of fluorophores and be in the same imaging conditions as for the described nanoclusters, the authors should use model systems composed of multimers of GFP vs. single GFP, immunolabeled with a GFP-binding monoclonal antibody. This will permit evaluation of the cluster signature obtained with DBSCAN analysis of STORM data for single vs. multimers of known stoichiometry. This would constitute an undisputable molecular stoichiometry ruler.

We appreciate the suggestions of the reviewer. Regarding the potential bias effect of the glass surface on the photophysics of the fluorophores we would like to clarify that the “calibration” for the number of blinking events per individual Ab on glass were performed on the same sample containing the cells that we image, so that we maintain exactly the same experimental and imaging conditions avoiding any potential artefacts. To our understanding this approach is more accurate than performing the calibration on glass substrates and then moving to samples containing the cells. This information is now contained in page 6 of the revised manuscript and in the materials and method section. Once the number of blinking events from individual Abs on glass within the DBSCAN radius are determined, one can then determine the number of localizations within the same DBSCAN radius on other parts of the sample. More localizations within the same DBSCAN radius basically means more molecules, and thus nanoclusters. This approach has been extensively used by other experts in the field as we properly acknowledge in our manuscript (Pageon et al, Mol. Cell. Biol 2016; Spiess et al. J. Cell Biol 2022).

Using anti-donkey antibodies to cross-link those STORM-specifically labelled Ab in order to artificially obtain clustered fluorophores, as suggested by the reviewer, is indeed a sound approach to retrieve signatures of clustering. Nevertheless, we have preferred not to use this approach because those artificially induced clusters would have very little resemblance to the real nanoclusters and would only allow us to validate the performance of DBSCAN for cluster recognition. As mentioned above, DBSCAN is a well-established algorithm and used by many different experts in the field and thus can be trusted by the community. Instead, and following the recommendation of the reviewer, we now provide results using an alternative cluster analysis algorithm (Voronoi tessellation) reaching similar conclusions regarding the existence of integrin and adaptor nanoclustering inside FAs.

Finally, the suggestion of using monomeric vs multimeric GFPs to determine the stoichiometry of the nanoclusters is highly appreciated. Indeed, we have used this approach in the past to identify nanoclustering of the chemokine receptor CXCR4 in living T cells (Mol. Cell 2018 and PNAS 2022). However, these experiments are best performed at sub-labelling conditions, which inherently underestimate the degree of nanoclustering. Combining GFPs with PALM to enable super-resolution is another approach but also subject to artefacts regarding the photo-conversion efficiency of GFPs as we reported earlier (Nature Methods 2017) and leading to underestimation of nanocluster stoichiometry.

In summary, providing nanocluster stoichiometry from single-molecule localisation images remains a major technical challenge and is highly sensitive to methodological assumptions. We have therefore focused here on providing robust evidence for the existence of integrin and adaptor nanoclustering, using three different superresolution approaches and two independent analytical methods for cluster determination.

Due to the surprising finding of the nanoclusters' "universality", it is imperative for the authors to validate the findings through complementary methodologies and analytical tools. This should involve replication of results using alternative super-resolution techniques (quantitative DNA-PAINT) and exploring different clustering algorithms (VoronoïTesselation) to ensure the robustness and reliability of the observations.

As already mentioned, we have now performed extensive DNA-PAINT to replicate most of our initial findings obtained by STORM, as requested by the reviewer. In addition, as a different super-resolution imaging strategy, we have also used STED microscopy to confirm the nanoclustering of integrins and some of the adaptors demonstrating now, by means of three different super-resolution techniques, that both integrins and their adaptors form nanoclusters of similar size and composition, regardless of whether they are inside or outside FAs. We have included these data as Figs. S3 and S4 and discussed the results in pages 8-9 of the main manuscript.

Regarding the use of an alternative analysis for the data, we have now used the Voronoi tessellation algorithm to re-analyse our STORM data, as requested by the reviewer. The results of the analysis, which render similar sizes and number of localizations as obtained by DBSCAN, are now included in Fig. S5 and mentioned in page 8 of the main manuscript.

Recommendations for the authors:

Reviewer #1 (Recommendations for the authors):

This work already, as is, provides significant and novel information on IAC.

The unexpectedly low spatial colocalisation of integrins with adaptor proteins might indeed be caused by the potentially quite long extension of talin upon force exposure and the ample zones of activity of IAC proteins, imaging the involved proteins in scale, as can be seen in Barnett and Goult (2022, doi: 10.3389/fncel.2022.1014629)? In super-resolution microscopy, this spatial separation might, in fact, become apparent. It would be interesting to see whether lowering the actomyosin contraction by different concentrations of blebbistatin lowers this separation. In general, it would also be interesting to understand whether lowering the forces can disrupt the nano-organisation and the strong separation of the two analysed integrin subtypes. It is true that nascent adhesion formation is force-independent, but maybe the forces play a role in establishing the particular integrin subtype nano-organisation within more mature IAC. I am also aware that a lot of work has already gone into the conclusion of this project.

We thank the reviewer for these thoughtful comments and suggestions. Our most recent preliminary data (not yet included in this manuscript) indeed indicate that the physical separation of integrin nanoclusters (and adaptors) inside focal adhesions (FAs) is force-dependent. We are currently reproducing these experiments using lipid bilayers of varying viscosities and controlled ligand density to explore how ligand mobility (i.e., equivalent to force exerted from the extracellular side) controls the degree of IAC nanoclustering and their spatial segregation in FAs. This approach is more amenable to super-resolution microscopy as lipid bilayers are quite thin and optically transparent, yet the experiments are still challenging, time-consuming, and thus ongoing.

To obtain a first hint as to whether forces might play a role in establishing the spatial distribution of the two different integrins within more mature IACs as the reviewer suggests, we have performed experiments at different seeding times (90 min, 3 hours and 24 hours). Our results show that even at earlier times (90 min), when a lower number of mature FAs are established, nanoclustering of integrins and main adaptors are similar to 24 hours. In contrast, and as suggested by the reviewer, the spatial distribution of the different subsets of integrin nanoclusters inside FAs is markedly different as a function of seeding time, with α5β1 nanocluster distribution being already established at 90 min, while αvβ3 nanocluster distribution appears rather random at earlier seeding times and progressively organizes reaching a well-defined lateral spacing at 24 hours of spreading time. As FA strengthening over time requires mechanical forces, and αvβ3 is preferentially involved in FA strengthening (Roca-Cusachs et al PNAS 2009), these data strongly suggest that forces play a differential role in the lateral distribution of both integrin nanoclusters over time. We have now included these data as a new Fig. 2 in the revised manuscript and discuss the results in the associated text (pages 9 and 10). We also mention in the discussion additional experiments, as suggested by the reviewer, to further substantiate this hypothesis.

Considering the high quality of the work and the new insight about the inner organisation of IAC, maybe the summary Figure 5 should be elaborated a bit, taking into account e.g. different lengths of extended talin proteins and also the various positions of vinculins on talin proteins (depending on opened cryptic binding sites), as well as the possibility that various actin filaments might be associated with single talins. What I mean is, the authors impressively demonstrate the complexity of IAC nano-organisation, which should be paid more tribute in the concluding figure. The quality of the figure should be adapted to the quality of the work.

We have adapted Figure 5 (now Figure 6) as suggested by the reviewer.

I would be curious to hear a bit more about the further speculations of the authors in the discussion, e.g., about why the integrin subunits are organised in this way. Why might the α5β1 be preferentially located in the periphery? What is the potential physiological relevance of this organisation? Is this organisation different in other cell types (have the authors looked at other cells)? Is the organisation lost in pathophysiological situations, such as cancer?

Although we do not know yet what drives the preferential location of α5β1 nanoclusters to the FA periphery, it is known that Kank2 also exhibits enrichment at the FA periphery, regulates talin activity and it is involved in the formation of α5β1-enriched fibrillar adhesions (Sun et al, Nature Cell Biol 2016). Thus, it is highly probable that α5β1 enrichment at the FA periphery is a necessary step for their translocation from mature FAs to fibrillar adhesions to then assemble fibronectin into the fibrillar networks as found and needed in connective tissues. Consistent with this idea, we have observed similar α5β1 distribution on other fibroblast cell lines (MEFS), which are the primary cells that produce fibrillar adhesions. Thus, α5β1 nanocluster distribution inside FAs might be physiologically important for the process of fibronectin fibrillogenesis.

Since it has been documented that tensin is important for fibronectin fibrillogenesis (Pankov et al J Cell Biol 2000) and more recently, it has been shown that tensin-3 interaction with talin drives the formation of fibronectin-associated fibrillar adhesions (Atherton et al, J Cell Biol 2022), we thought to investigate the spatial distribution of tensin-3 and its relationship with α5β1 inside FAs by means of dual colour super-resolution STED microscopy. Interestingly, our initial data on HFF cells seeded for 24 hours show both enrichment of tensin-3 and α5β1 nanoclusters at the edges of mature FAs, supporting our working hypothesis that α5β1 enrichment at the FA periphery serves to translocate α5β1 integrins from FAs to fibrillar adhesions, probably in a talin-tensin-dependent manner. While these initial data are quite exciting, many more experiments that include simultaneous super-resolution mapping of α5β1, talin and tensin in mature FAs are required to fully validate our hypothesis. Yet, because of their relevance we consider it appropriate to include these data as Fig. S8 and discussing their potential implications in pages 22 and 23 of the revised manuscript.

Reviewer #2 (Recommendations for the authors):

(1) Perform control experiments to confirm that the nanocluster size/composition is not affected by the antibodies used.

As explained in the response to the public reviews, antibody labelling has always been performed after cell fixation, precluding potential cross-linking artefacts due to protein mobility and avoiding unwanted receptor activation. In addition, we have performed super-resolution imaging using DNA-PAINT as a different imaging strategy. In this case, the DNA docking site is site-specifically coupled to one camelid single-domain Ab (sdAB), having a much smaller size as compared to a secondary Ab, reducing therefore linkage error and increasing the accessibility of primary Ab-labelled proteins. As can be observed in new Fig S4, no differences in terms of nanocluster sizes and/or compositions were observed for any of the proteins investigated using DNA-PAINT as compared to our initial STORM data. These control experiments thus rule out any potential artefacts introduced by the secondary Ab. Finally, we would like to highlight that our results on the nanoclustering of integrins in terms of their size and number of localizations is consistent with previous results obtained by other groups around the world using similar labelling protocols as us (Spies et al, J. Cell Biol 2022), or relying on halo-tag strategies, as suggested by the reviewer (see Fujiwara et al, J. Cell Biol. 2023). The consistency of these results amongst different groups gives us further confidence that the nanocluster size/composition are not affected by the antibodies used.

(2) Extend the study by inspecting a set of integrin adapter proteins for their association with inactive integrins, focusing on adapters associated with the maintenance of the inactive state. Possible candidates would be tensin and filamin, for example.

We thank the reviewer for the suggestion and have now performed dual-colour super-resolution STED microscopy of α5β1 and tensin-3 on HFF cells seeded for 24 hours. Interestingly, instead of being an integrin inactivator, we found that tensin-3 is also highly enriched at the FA periphery where a large fraction of active β1 integrins are located, suggesting that at these particular regions, active β1 could be either engaged to talin (as shown in our original data) or to tensin-3 (our new data shown in Fig. S8). These results might be surprising at first, since tensin competes with talin for the same binding site to the cytoplasmic β-tail of integrins, and thus believed to act as integrin inactivator, as the reviewer indicates. Nevertheless, recent data has shown that tensin is capable to activate integrins (in particular if β1 is phosphorylated) by interacting with the actin cytoskeleton, providing mechanical coupling for integrin activation (Georgiadou & Ivaska, Trends Cell Biol. 2017). We have now included these new data as Fig. S8 in the revised manuscript. Additional experiments, which in our opinion fall outside of the scope of this work, would be necessary to identify other potential integrin inactivator partners, but certainly a topic of future interest to our group.

(3) While the methods are described in sufficient detail, it is important to ask if the findings are based on sufficient data. Table S5 provides detailed information about the number of samples studied, and it appears that only small numbers of samples were investigated for certain protein pairs. This should be discussed, and perhaps more data should be obtained to strengthen the data.

We have now performed additional experiments using DNA-PAINT as alternative super-resolution imaging technique (as also requested by reviewer 3) which adds additional data to the whole manuscript.

Version 2

Abstract

α5β1 and αvβ3 integrins are core components of focal adhesions (FAs) involved in cell attachment, migration, and mechanobiology-dependent processes. Recent works indicate that both integrins organize in nanoclusters inside FAs, with sub-populations of active and inactive β1 nanoclusters. However, whether both integrins work in concert or their activities are spatially regulated is not fully understood. Using dual-color super-resolution microscopy (STORM, STED and DNA-PAINT) we show that integrins α5β1 and αvβ3 and their main adaptor proteins exhibit similar functional nanoscale segregation across all adhesion proteins examined, independent of FA maturation state. Notably, both integrins never mix at the nanoscale indicating that their functions might be spatially regulated. We find a nearly 1:1 relationship between active integrin and adaptor nanoclusters suggesting that coordinated integrin activation occurs via the concurrent engagement of adaptor nanoclusters. Interestingly, α5β1 nanoclusters preferentially localize at the FA periphery near adaptor nanoclusters, establishing regions of multi-nanocluster enrichment, whereas αvβ3 nanoclusters uniformly distribute throughout FAs. Overall, our results show that adhesion proteins are arranged as modular nanoscale units that distinctively organize inside FAs to spatially regulate integrin activation and function.

Introduction

Adhesion complexes (ACs) in adherent cells, such as focal adhesions (FAs) and fibrillar adhesions (fAs) are places for cell attachment, mechanosensing and mechanotransduction (1–4). These structures are composed by different integrin receptors that bind to their ligands on the extracellular matrix (ECM), and by a myriad of intracellular adaptors, proteins that dynamically connect integrins to the actin cytoskeleton machinery relaying the transmission of mechanical forces in a bidirectional manner (5–7). Studies over the last two decades have provided a wealth of information focused on understanding these complexes, the proteins they contain, their role in cellular processes, such as mechanosensing, migration and proliferation, as well as how influential their regulation is in developmental and pathogenic processes (8–15).

Integrins α5β1 and αvβ3 are transmembrane heterodimeric proteins critical for the formation of ACs. Both α5β1 and αvβ3 bind to the RGD peptide motif of certain ECM components, such as fibronectin (16–18), and are linked to the cell’s cytoskeleton via many of the same adaptor proteins (19–22). The apparent redundancy of different integrins binding to the same ligand and same adaptor proteins is intriguing and has been the subject of intense research over the past years. Earlier studies on primary fibroblasts showed that these integrins segregate in different adhesion structures, indicating different roles, with α5β1 being enriched in fAs (23), adhesive sites at central regions of the cell, and αvβ3 being confined to focal complexes at the base of cellular protrusions, and to FAs, which anchor contractile stress fibers at the periphery of the cell (24). However, more recent super-resolution microscopy studies showed that both types of integrins are simultaneously present inside FAs (25). Within FAs, α5β1 has been proposed as the integrin responsible for adhesion strength while αvβ3 enables mechanotransduction, as cells exposed to α5β1 blocking antibody exhibited weaker attachment than those whose αvβ3 integrins were blocked (5). These results are also aligned to in vitro reconstitution experiments, that have assessed the catch-bond behavior of both integrins with fibronectin. In this type of bond, the lifetime of the bond increases as the applied force increases, up to a certain threshold, beyond which the bond ruptures. The rupture force for α5β1 binding to its ligand has been found to be higher (∼10–30 pN) than that of αvβ3 (∼5–25 pN) (26, 27). Nevertheless, single molecule tracking data indicated that αvβ3 integrins are mainly stationary inside FAs and more stably bound to the ligand, favoring adhesion as its main role (25). It has been also shown that α5β1 is involved in Rho activation during late spreading, enhancing RhoA-mediated actomyosin contractility (24, 28), supporting the idea that α5β1 is involved in force-mediated adhesion reinforcement. In a tour-de-force effort, Fässler and colleagues found that only the simultaneous expression of both integrins enabled cells to sense the rigidity of the ECM and to respond by reinforcing the actomyosin structure (29). These results, supported by others (30), have led to the recent notion that both integrins work in a synergistic manner, exhibiting cooperativity during cell adhesion, migration and mechanosensing, although the mechanisms are far from being fully resolved (30).

The highly dynamic conformational states exhibited by integrins together with their lateral and axial nanoscale organization might play important roles regulating the differential or synergistic functions of different integrins inside FAs. Indeed, it has been shown that integrins can rapidly transition between different conformational states that relate to their activation and binding capacity to their ligands (25, 31, 32). Moreover, in addition to the nanoscale axial segregation of different AC components inside FAs (33), it has been reported that integrins as well as different adaptor proteins also segregate laterally into small nanoclusters inside FAs (34–37). Recent super-resolution studies have further shown that β1 integrins segregate in both active and inactive nanoclusters inside FAs (36) and single-molecule tension measurements uncovered distinct sub-populations of load bearing integrins and a high fraction of unengaged integrins (38, 39). Together, these results suggest that at any given time only a fraction of integrins might be active inside FAs and engaged with their adaptor partners. Such a dynamic scenario would allow integrins to differentially interact with their environment and possibly diversify their functions. As α5β1 and αvβ3 have different activation states and are both contained in FAs, it is plausible that their spatial regulation and activation state inside FAs provides a physical mechanism to tune their functions. However, such studies have not been reported yet.

Here, we combined dual-color super-resolution microscopy together with robust analytical tools to study a number of endogenous FA proteins on human fibroblast cells. Specifically, we imaged one of the two integrins (α5β1/αvβ3), their full population or active conformations, with each of three key adaptors proteins, paxillin, talin and vinculin. Our results show that each of these proteins form segregated nanoclusters of remarkably similar size and molecular density inside FAs, and interestingly, also on the membrane outside of ACs. Moreover, we find sub-populations of active and inactive β1 and β3 nanoclusters inside FAs, which spatially segregate from each other and provide evidence for a nearly 1:1 relationship between talin, vinculin and active integrin nanoclusters. Interestingly, we find that αvβ3 nanoclusters organize randomly inside FAs regardless of their activation state, whereas α5β1 and in particular the active conformation of β1 nanoclusters preferentially organize at the edges of FAs. Overall, we propose that the nano- and meso-scale organization of α5β1 and αvβ3 integrins and their distinct lateral segregation and activation state are key factors in the differentiation of their roles within FAs.

Results

Proteins of the ACs organize as universal nanoclusters across regions of the basal membrane

We employed dual-color super-resolution microscopy based on stochastic optical reconstruction microscopy (STORM) to determine how adhesion proteins (integrins and key adaptor proteins) spatially organize at the nanoscale. Human fibroblasts (HFF-1 cells) were seeded on fibronectin (FN)-coated glass slides for 24 h, after which the cells were processed for immunofluorescence. One integrin (either the total population of α5β1 or αvβ3, or the active population specific to the β1 or β3 integrin subunit) was labeled together with one adaptor protein (paxillin, C terminus of talin-1, or vinculin) and then imaged using STORM. We note that while the pan antibodies used to target the full population of integrins are specific to α5β1 or αvβ3, (with the anti-α5 antibody binding the α5 subunit, which only dimerizes with β1, and the anti-αvβ3 targeting this specific heterodimer), the active antibodies target the β subunits. These β subunits can dimerize with other α subunits, but with our experimental conditions, where FN is the most abundant ligand (see Materials & Methods), the primary integrins being probed are the heterodimers, α5β1 or αvβ3. Given the huge number of adaptor proteins that have been reported to interact with integrins (19, 20), we chose for paxillin, talin and vinculin since they belong to the core-components of the adhesome (21) and crucially locate at distinct axial layers of functional activity, i.e., paxillin at the signaling layer, while talin and vinculin locate at the force-transduction layer (33).

STORM super-resolution images qualitatively revealed an inhomogeneous distribution for all the proteins examined, forming small and laterally segregated nanoclusters inside FAs (Fig. 1A–D). To quantitatively substantiate these observations, we first used a mask to classify all the single molecule localization events according to their location on the membrane, i.e., FAs, fAs or outside ACs (see Materials & Methods, Fig. S1A-F). To identify whether multiple localization events correspond to nanoclusters we used a density-based spatial clustering of applications with noise (DBSCAN) algorithm (40), which has been extensively used for quantifying protein nanoclustering from single molecule localization images (41–44), including β1 integrins inside FAs (36, 45). Since STORM data consist of lists of single-molecule localizations and each labeled protein can be localized multiple times due to fluorophore blinking (46) resulting in pseudoclusters (i.e., non-biological artefactual clusters (41)), we first assessed the number of localizations detected due to the multiple blinking events of a single fluorophore. For this, we followed an approach similar to that reported by Spiess et al. to generate distributions of the number of localizations obtained from isolated fluorescent spots on glass next to the cells likely corresponding to individual antibodies (Abs), for both channels (Fig. 1E, Materials & Methods). We note that providing quantification on the FN-coated glass substrate next to the cells guarantees identical experimental and imaging conditions, thus avoiding potential differences in blinking photophysics as compared to calibration measurements performed solely on glass (i.e., without the cells). By following this approach, one can determine the number of localizations corresponding to individual Abs and infer whether the clustered localizations observed correspond to individual or multiple proteins. The median values of the distributions for individual Abs were 5 localizations for Alexa 405-Alexa 647-conjugated Abs (Channel 1) and 3 localizations for Cy3–Alexa 647-conjugated Abs (Channel 2) (Fig. 1E). Moreover, to further minimize the effect of fluorophore blinking on cluster identification, we imposed an additional threshold of 10 localizations as commonly used in the field (36, 45). Thus, any cluster identified by DBSCAN with 10 or more localizations was considered as a cluster containing more than one protein (see Materials & Methods). DBSCAN classified between 70% to 80% of the localizations of each of the proteins as belonging to nanoclusters, regardless of whether they were located inside FAs or on regions of the membrane outside ACs (Fig. S1G-I). Notably, the number of localizations per nanocluster and their size were similar across the different adhesion proteins (integrins and adaptors) inside FAs, with a median of ∼20 localizations per nanocluster, likely corresponding to a few copies of the protein, and a diameter of ∼ 40–45 nm (Fig. 1F, G).

Proteins of FAs organize as universal nanoclusters of similar size and containing a similar number of proteins.

(A) Representative dual-color STORM image of an HFF-1 cell fixed after 24 h on FN and labeled for α5β1 (magenta) and paxillin (green). Scale bar: 5 μm. (B) Representative images of all the adhesions identified for α5β1 with paxillin (left), talin (middle) and vinculin (right), with zoomed-in regions shown below each panel. Scale bar top row: 5 μm, bottom row: 1 μm. (C) Representative images of all the adhesions identified for αvβ3 with paxillin (left), talin (middle) and vinculin (right), with zoomed-in regions shown below each panel. Scale bar top row: 5 μm, bottom row: 1 μm. (D) Representative images all the adhesions identified for of active β1 (act β1) with paxillin (left), active β3 (act β3) with paxillin (right), with zoomed-in regions shown below each panel. Scale bar top row: 5 μm, bottom row: 1 μm. (E) Distribution for the number of localizations per secondary antibody on glass next to cell regions for both of the secondary antibodies used in this study, goat anti-mouse Alexa 405–Alexa 647 (Ch1) and goat anti-rabbit Cy3–Alexa 647 (Ch2). The median number of localizations per spot corresponds to: 5 for Alexa 405-Alexa 647-conjugated antibodies and 3 for Cy3–Alexa 647-conjugated antibodies (see Materials & Methods). (F, G) Box-and-whisker plots showing the number of localizations per nanocluster (F) and the nanocluster diameter (G) for the proteins in this study. The box represents the range from the 25th to the 75th percentiles and whiskers from 10th to 90th percentiles, with the median indicated by a horizontal line. Points outside these whiskers correspond to outliers. Individual points (points within the whiskers not shown) correspond to the median value over all nanoclusters for each cell (for N values, i.e., number of independent experiments, see SI Table S5). No statistical differences were found using the multi comparison one-way ANOVA test, where not significant (ns) corresponds to p>0.05.

Similar results in terms of diameter and number of localizations per nanocluster were obtained when performing the analysis outside ACs (Fig. S2). In addition, the nanocluster characteristics were maintained when specifically labeling the active integrin (β1 or β3) populations (Fig. 1 D, F, G). These results are in good quantitative agreement with previous reports using STORM showing that both active and inactive integrin β1 nanoclusters have a median size of 40 nm and 35 nm, respectively, as well as comparable number of single molecule localizations per nanocluster (36). Notably, our results provide new insights on the integrin αvβ3 and the main adaptor proteins, paxillin, talin and vinculin, by indicating that the organization into laterally segregated nanoclusters of similar size and number of molecules is maintained across different protein types. Moreover, such nanoscale organization appears to be independent of whether these adhesion proteins partition inside FAs or outside ACs (Fig. S2).

Given the remarkable similarity in terms of nanoclustering size and number of localizations amongst the different proteins investigated and regardless of their location (i.e., inside or outside FAs), we performed super-resolution imaging using two additional and different methodologies, stimulated emission depletion (STED) microscopy (47, 48) and DNA-points accumulation for imaging in nanoscale topography (DNA-PAINT) (49). Because STED and DNA-PAINT rely on different physical mechanisms (fluorescence depletion via stimulated emission depletion in STED, and kinetic binding of short dye-labeled oligonucleotides to their complementary target strands in the case of DNA-PAINT) as well as different technical implementations, they constitute excellent controls to validate the results obtained by STORM. Moreover, albeit DNA-PAINT is a single molecule-based localization method similar to STORM, it is not influenced by fluorophore blinking (50), providing an additional control to our STORM results. Both STED and DNA-PAINT images confirmed nanoscale clustering of the proteins investigated, both inside and outside FAs (Fig. S3, Supplementary Text 1, and Fig. S4, Supplementary Text 2), robustly validating our STORM results.

To further exclude any potential bias in quantifying the extent of nanoclustering as obtained by DBSCAN, we implemented an algorithm based on Voronoi-tessellation (41) and re-analyzed all the STORM data (Supplementary Text 3). This analysis showed similar nanoclustering size and number of localizations per nanocluster as compared to DBSCAN (Fig. S5). Altogether, our experimental results obtained using three distinct super-resolution modalities and by analyzing the images using two different clustering algorithms render comparable sizes in terms of nanoclustering for all the proteins investigated, indicating that proteins of ACs organize as universal nanoclusters of similar size and number of proteins, regardless of whether these nanoclusters are located inside or outside FAs. Nevertheless, we note that quantifying the actual stoichiometry of the nanoclusters, i.e., number of proteins per nanocluster based on fluorescence strictly requires a 1:1 labeling between the protein and the fluorophore which is highly challenging, in particular when employing super-resolution approaches (51–53). Therefore, our relative comparison of nanoclustering among the different proteins investigated is exclusively based on the number of single molecule localizations contained in each nanocluster (both for STORM and DNA-PAINT) which is a fair approach since we always use the same reporter fluorophore and maintain similar excitation conditions throughout our experiments.

α5β1 and αvβ3 integrin nanoclusters differentially organize inside FAs

Our results so far indicate that the two major integrins involved in FAs form nanoclusters of similar size and molecular content. To enquire whether they act in concert inside FAs by associating at the nanoscale, we performed dual color STORM imaging of the total populations of α5β1 and αvβ3 integrins. For this, we used a rabbit anti-α5 antibody (Table S2) to enable simultaneous labeling with the mouse anti-αvβ3 antibody that we used throughout this study. Visual inspection of the images already suggests that α5β1 and αvβ3 do not physically overlap at the nanoscale (Fig. 2A). To robustly quantify these observations, we calculated the edge-to-edge distance between the nearest-neighbor (EED-NN) integrin nanoclusters (i.e., closest α5β1 to αvβ3 and closest αvβ3 to α5β1) (see Materials & Methods). Here, EED-NN values below zero correspond to physically overlapping α5β1 and αvβ3 nanoclusters, EED-NN values above zero to spatially segregated nanoclusters, and EED-NN = 0 corresponds to contiguous α5β1 and αvβ3 nanoclusters, i.e., next to each other. Quantification of the images confirmed that α5β1 and αvβ3 do not intermix at the nanoscale as the percentage of EED-NN values below zero is negligible and comparable to values obtained from simulations of random organization of both integrins (Fig. 2B). Thus, α5β1 and αvβ3 integrins organize as distinct nanoclusters that spatially segregate from each other inside FAs.

α5β1 and αvβ3 integrin nanoclusters differentially organize inside FAs.

(A) Representative dual-color STORM image of the total populations of α5β1 (magenta) and αvβ3 (green) integrins inside FAs. The panel on the far-right side corresponds to a zoomed-in region of a single FA (white box) showing spatially segregated α5β1 and αvβ3 nanoclusters. Scale bars: 2 μm, and 500 nm for the zoomed-in region. (B) Histogram distribution for edge-to-edge distance between nearest neighbors nanoclusters (EED-NN), from α5β1 to αvβ3 (A, green) and αvβ3 to α5β1 (B, blue). (C-D) Nanocluster properties inside FAs for the integrins α5β1 (green), and αvβ3 (blue) as a function of seeding time. (C) show the number of localizations per nanocluster;(D) shows the nanocluster diameter. Only the pairs with statistical differences are marked. *, p<0.05; **, p<0.01; ***, p<0.001; one-way ANOVA. (E-H) Histograms of the experimental NND distributions for the α5β1 (green) (E) and αvβ3 (blue) (G) integrins in FAs, for the designated cell spreading times. Histograms for the random simulations have been omitted for the sake of visual clarity. (F,H) ΔNNDH-plots for the subtraction histograms for simulated data from experimental for the α5β1 (F) and αvβ3 (H) integrins in FAs at the three different seeding times (90 min, 3 h and 24 h).

Since FA formation is a dynamic process where mechanical forces and FA maturation increase over several hours, we thought to investigate whether the nanoscale organization and distribution of both integrins inside FAs are sensitive to FA maturation. For this, we performed STORM imaging as described above on cells seeded for 90 min and 3 hours and quantified the number of localizations per nanocluster (Fig. 2C) and nanocluster diameter (Fig. 2D) for both integrins and, compared the results to 24 hours cell seeding. No differences in nanocluster properties were observed for any of the two integrins suggesting that integrin nanoclustering is independent of spreading times and FA maturation. To then investigate how integrin nanoclusters are distributed inside FAs, we performed nearest-neighbor distribution (NND) analysis for both integrins at different cell seeding times (Fig. 2E, G). Since the NNDs are influenced by the nanocluster densities, we compared the experimentally obtained NNDs with those obtained from simulations of randomly distributed nanoclusters considering the experimentally obtained nanocluster density. We then calculated the difference between experimental and simulated data for each NND value, i.e., ΔNND for both integrins at different cell spreading times (Fig. 2F, H). Interestingly, our analysis revealed a clear deviation from random organization inside FAs, with both integrin nanoclusters being segregated from each other at around 55 nm and enriched at distances ∼ 100-150 nm. Strikingly, this physical segregation and enrichment depended on cell spreading time and it was different for both integrins: α5β1 segregation and enrichment was already established at earlier spreading times (90 min) (Fig. 2F), while αvβ3 nanoclusters organized rather randomly inside FAs at 90 min and progressively established their characteristic spatial segregation at ∼ 55 nm and enrichment at ∼ 120 nm at 24 hours of cell spreading (Fig. 2H). Overall, these results indicate that while integrin nanocluster properties are preserved and maintained regardless of cell spreading times and FA maturation, their distribution and organization inside FAs depends on FA maturation, with αvβ3 lateral ordering being progressively reached at later stages of cell adhesion.

Lateral organization of adaptors and integrin nanoclusters inside FAs

We next sought to investigate how the different adaptor proteins laterally organize inside FAs beyond their nanoclustering. For this, we first quantified the density of nanoclusters per unit area of the membrane, within FAs (Fig. 3A) and outside ACs (Fig. 3B). As expected, the density of nanoclusters was larger in FAs as compared to outside adhesions (∼5–10-fold increase) (Fig. 3A, B). Adaptor proteins, such as talin or vinculin, are shared between α5β1 and αvβ3 integrins (21), therefore we compared the density of the full population of integrin nanoclusters (∼36 µm-2 in FAs) to the density of adaptor protein nanoclusters (∼10–12 µm-2 for talin and vinculin, ∼20 µm-2 for paxillin in FAs). These data reveal a significantly higher (∼2–4 fold) density of integrin nanoclusters in comparison to their adaptors and, consequently, a nanocluster mismatch between the number of integrins and adaptor nanoclusters, regardless of whether they partition inside or outside FAs (Fig 3A, B). Since the active integrin conformation is the one engaged with the adaptors (54, 55), we also quantified the density of active integrin nanoclusters (active β1 and active β3) per unit area inside FAs. Our results showed that only ∼ 30% of the total α5β1 nanocluster population was in an active conformation, and this percentage was even lower for β3 nanoclusters (∼ 20%) (Fig. 3A), indicating that at any given time there is a large percentage of inactive integrin nanoclusters inside FAs. We note that these percentages might be even lower considering that our Abs for mapping the active conformations are not exclusive for α5β1 and αvβ3, but rather to the total active β1 and β3 sub-populations.

Lateral organization of integrin nanoclusters and adaptor proteins inside FAs.

(A, B) Mean number of nanoclusters per unit area inside FAs (A) and on the cell membrane outside adhesions (B), with α5β1 and actβ1 in light and dark green, respectively; and αvβ3 and actβ3 in light and dark blue, respectively. The number of integrins nanoclusters per unit area are shown as stacked graphs to allow comparison between the total number of integrin nanoclusters (α5β1 and αvβ3) and the adaptor proteins. The bars show the mean ± st. dev of the distribution. Statistical significance between data of all proteins was tested using a multi comparison one-way ANOVA, assuming Gaussian distribution of points. The comparison between α5β1 and αvβ3, (actβ1 and actβ3) was done with an unpaired t-test, where ns = not significant, p>0.05; *, p<0.05; **, p<0.01; ***, p<0.001. The insets show representative images of nanoclusters for α5β1 (magenta) and paxillin (green) in FAs (A) and outside adhesions (B). Scale bar: 1 μm. (C) Histogram of the EED-NN values between α5β1 and the different adaptors (green) and, between αvβ3 and the adaptors (blue). The black line in each plot corresponds to the histogram for the simulated data sets. The dashed red line at zero corresponds to nanoclusters contiguous to each other. Percentage of overlapping nanoclusters: α5β1-paxillin: exp = 28.4%, sim = 24.4%; α5β1-talin: exp = 6.7%, sim = 8.1%; α5β1-vinculin: exp = 6.0%, sim = 7.7%; αvβ3 -paxillin: exp =31.7%, sim=29.2%; αvβ3 -talin: exp =15.7%, sim=16.5%; αvβ3-vinculin: exp =14.9%, sim=17.4%. (D) Histogram of the NND values between nanoclusters of the same protein. The black line in each plot corresponds to the histogram for the simulated data sets. Bin width is 10 nm.

Notably, the total density of active β1 and β3 integrin nanoclusters matches that of talin and vinculin nanoclusters, and it is only slightly outnumbered by paxillin nanoclusters inside FAs (Fig. 3A), suggesting that these active integrins might be fully engaged to their adaptors. The modest percentage of active β1 integrin nanoclusters observed in our data (∼ 30%, Fig. 3A) is in close agreement with a recent report from Spiess et al, obtained using a different super-resolution technique (36) and consistent with the high percentage of low-force bearing integrins measured by using single-molecule force sensors (39). In addition to the α5β1 results, our data also reveal the existence of a sub-population of active β3 nanoclusters inside FAs (Fig. 3A). Together, these results strengthen the hypothesis proposed by Spiess et al. of a coordinated mechanism that laterally segregates active and inactive nanoclusters inside FAs (36). Moreover, the differences in expression levels, activation states, and lateral nano- and lateral scale segregation of α5β1 and αvβ3 nanoclusters inside FAs suggest the existence of a mechanism that defines the distinct functional roles of these two integrins inside FAs.

To get a better understanding on the lateral organization of both types of integrins in relation to their main adaptor proteins, we calculated the EED-NN between integrin nanoclusters and each of the adaptors. Although visual inspection of the dual color STORM images already suggested that integrin nanoclusters are also physically segregated from their adaptor partners (Fig. 1B, C) we used the EED-NN analysis to estimate the degree of colocalization (i.e., spatial overlap) between different nanoclusters. Our results confirm that the large majority of α5β1 and αvβ3 nanoclusters are laterally segregated from their adaptor partners (Fig. 3C, Fig. S6). Their spatial segregation is essentially determined by the density of nanoclusters within the FAs, as similar EED-NN distributions were obtained from simulations of particles uniformly distributed inside the experimentally obtained FAs, using identical particle densities as to the ones obtained experimentally for the integrin nanoclusters and the different adaptors (Materials & Methods) (Fig. 3C, black lines). Considering the canonical view that integrins inside FAs are bound to paxillin and engaged to talin and vinculin (55, 56), these results might be surprising at first sight, as one would expect a much higher fraction of EED-NN values below zero. Nevertheless, our data showed only a small percentage of active integrin nanoclusters as compared to their total populations (Fig. 3A) so that spatial overlap with talin and/or vinculin would be at best 30% for α5β1 and 20% for αvβ3 nanoclusters, which under our experimental conditions might be undistinguishable from a random overlap. Notably, using different super-resolution microscopy techniques, others have also found negligible spatial colocalization of integrins with these adaptor proteins (36, 37). Moreover, Dunn and co-workers also found that most ligand-bound integrins inside FAs exist in a minimally tensioned state (< 2 pN) and not dependent on the actin cytoskeleton (39). The results from the Dunn lab are fully in line with the lack of colocalization of integrins to talin and vinculin as we determined here, since these two adaptors mediate the linkage of integrins to the actin cytoskeleton (55, 57). Nevertheless, we would like to mention that we labeled talin at its actin-binding C-terminal region, which in a 2D projection can appear to be at a distance of 100 nm or more from the N-terminal integrin-binding site (58), so lack of spatial overlap does not necessarily indicate lack of direct engagement.

It has been recently proposed that adhesion proteins including integrins organize in a hierarchical fashion inside FAs, forming nanoclusters similar to the ones reported here, that further enrich into compartmentalized regions around 300 nm in size (37). To enquire whether such hierarchical organization is present in our data, we performed NND analysis between nanoclusters of the same proteins (Fig. 3D). In addition, we implemented simulations of uniformly distributed particles inside the experimental FA regions, using the same experimental nanoclusters densities, and made the corresponding NND analysis (Fig. 3D). Our results show that the global organization of all the adhesion protein nanoclusters investigated is highly uniform inside FAs, as the experimental NND distributions were indistinguishable from those of randomly distributed particles (Fig. 3D). In contrast to the recent findings of the Kusumi lab (37), we did not observe any evidence of spatial enrichment of protein nanoclusters inside FAs. Instead, the observed characteristic separation between nanoclusters appears to be solely determined by their density inside FAs.

Integrin α5β1 but not αvβ3 nanoclusters are enriched on FA edges

Our previous NND analysis rendered a uniform inter-cluster organization of all adhesion protein nanoclusters inside FAs. Yet, visual inspection of the dual-color STORM images suggests that a significant fraction of α5β1 nanoclusters preferentially localize on FA edges (Fig. 1B). To quantify these observations, we calculated the shortest-path distance from each α5β1 nanocluster to the edge of its corresponding FA, i.e., its distance-to-FA-edge, and generated a histogram of these values (Fig. 4A). In addition, we generated a similar histogram from simulations of uniformly distributed α5β1 nanoclusters on FAs (see Materials & Methods). In general, if the nanoclusters were uniformly distributed throughout the FA, one would expect that the probability of finding a nanocluster at a specific distance to the FA edge decreases with increasing distance. This occurs because there is a larger area along the perimeter of any given shape and thus a higher number of nanoclusters (e.g., for a circular shape, the decrease is linear with the distance to the FA edge). We found that our experimental and simulated data indeed exhibit this general trend, although, notably, a higher percentage of α5β1 nanoclusters were detected at distances below 100 nm from the FA edges, as compared to a uniform distribution (Fig. 4A and inset), indicative of their preference to locate closer to the FA edges. Similar trend was also obtained for talin, and to a lesser extent for paxillin, whereas the distributions of αvβ3 and vinculin were indistinguishable from random (Fig. S7A).

Integrin α5β1 nanoclusters are preferentially distributed on the edges of FAs.

(A) Distribution of distances from the center of mass (CoM) position of each individual α5β1 nanocluster to its closest FA edge obtained from DBSCAN-analyzed STORM images. The histogram represents the frequency distribution over all cells. Green lines correspond to the experimental data (Exp) and the simulated uniform distribution (Sim) is shown as a black histogram. Bin width is 10 nm. The inset shows the distribution for the first 250 nm. (B) Edge Proximity Factor (EPF) per cell for each of the proteins investigated. Data are shown as median values per individual cell (color dots). The mean ± st. dev. over all the cells are superimposed as black lines. Statistical significance between experimental and simulated data was tested using a paired t-test, where ns = not significant, p>0.05; *, p<0.05; **, p<0.01; ***, p<0.001. (C, D) Variation of the nanocluster distances from the FA edges with respect to random organization, as function of the distance to the FA edges (d2e) for α5β1 and αvβ3 integrin nanoclusters (C) and for the adaptor proteins (D). (E) Percentage of nanoclusters proximal to FA edges (d2e<100nm) for the total and active β1 nanocluster population (color dots) together with results from simulations of random organization (black dots). Each dot corresponds to the median value per cell, each cell containing multiple FAs. The mean ± st. dev. over all the cells is superimposed as red lines. Statistical significance between experimental and simulated data was tested using a paired t-test, and to test the significance between α5β1 and actβ1 the unpaired t-test was used, assuming a Gaussian distribution of data, where ns = not significant, p>0.05; *, p<0.05; **, p<0.01; ***, p<0.001.

To better assess the extent of nanocluster enrichment at FA edges at the single cell level, we computed the median values of the experimentally obtained distance-to-FA-edge distributions per cell (over multiple FAs) and those corresponding to the simulated data. We then defined the edge-proximity-factor (EPF) as the relative difference between experimental and simulated distributions, where negative EPF values correspond to preferential nanocluster enrichment at FA edges, positive EPF values to nanocluster depletion at FA edges, and EPF = 0 corresponds to random location of nanoclusters with respect to the FA edges (Fig. 4B). Moreover, to statistically determine whether the distribution of EPF values was different from zero (the EPF value expected from a random distance-to-the-FA-edge distribution) we ran a paired Student’s t-test for each protein comparing the experimental with the corresponding simulated distributions. This analysis showed that most proteins exhibit some degree of enrichment at the edges of the FAs (i.e., negative EPF values), although their magnitude was markedly different amongst them (Fig. 4B). The highest enrichments at the FA edges were found for the full population of α5β1 nanoclusters and the active β1 sub-population (∼30% closer to the FA edge than if uniformly distributed) and for talin, whereas the enrichment at the FA edges of the total αvβ3 population and paxillin was much more modest. Notably, the active β3 sub-population and vinculin showed no preferential enrichment at the FA edges (Fig. 4B).

To establish the characteristic distances at which nanocluster enrichment occurs with respect to the FA edges, we next computed the difference between the experimental and simulated distributions, i.e., the variation from random, as a function of the distance to the FA edges, for each of the examined proteins (Fig. 4C, D). Here, positive values correspond to local nanocluster enrichment at a given distance to the FA edge, whereas negative values indicate nanocluster depletion. Our results show a clear difference between both integrins, with α5β1 nanoclusters being enriched close to the edge of FAs (Fig. 4C, continuous green line) whereas αvβ3 nanoclusters showed a rather uniform distribution, regardless of their distance to the FA edges (Fig. 4C, continuous blue line). Moreover, the preferential enrichment of β1 but not of β3 nanoclusters at the FA edges was retained when specifically analyzing the active integrin conformations (Fig. 4C, dashed green and blue lines, respectively).

We next investigated how adaptor protein nanoclusters distribute with respect to the FA edge using a similar approach. We found that both talin and paxillin nanoclusters showed enrichment in a region between 50–150 nm from the FA edges, although somewhat less pronounced as compared to α5β1 nanoclusters (Fig. 4D), whereas vinculin nanoclusters showed no preference to reside proximal to the FA edges. Finally, we extracted the percentage of the total α5β1 and active β1 nanoclusters that reside within distances shorter than 100 nm from the FA edges and compared the results to those from simulations of random organization (Fig. 4E). Interestingly, a significantly larger percentage of active β1 nanoclusters as compared to the total α5β1 nanocluster population and to the simulations, locate at the periphery of FAs (Fig. 4E).

Together, these results strongly indicate that integrin nanoclusters spatially organize in a rather distinct fashion within FAs, with α5β1 nanoclusters localizing preferentially at the edges of the FAs and αvβ3 nanoclusters distributing randomly inside FAs. Moreover, the preferential enrichment of active β1 nanoclusters together with talin at the FA edges suggests that β1 activation is spatially regulated inside FAs.

Active β1 nanoclusters form multiprotein hubs at FA edges

Since activation of integrins involves engagement with their adaptors paxillin, talin and vinculin, we next asked whether the active β1 nanoclusters preferentially located at the FAs periphery also reside in proximity to their adaptor proteins. For this we first obtained the inter-nanocluster NND (iNND) distributions between active β1 nanoclusters and each of the adaptor nanoclusters (Fig. S7B). We calculated the median of each iNND distribution and used this value as a threshold to classify pairs of active β1-adaptor nanoclusters as being close (i.e., iNND < median) or far (iNND > median) from each other. Next, we separately computed the distance-to-the-FA edge for the active β1 nanoclusters that were found close or far from their partner and compared the results to simulated data (Fig. 5A and inset). Having these plots, we then calculated the variation from random (i.e., difference between experimental data and simulations of random organization) as a function of the distance to the FA edge, for both active β1 nanoclusters close to their partner and active β1 nanoclusters far from their partner (Fig. 5B and Fig. S7C). Surprisingly, this analysis shows that active β1 nanoclusters preferentially locate the FA edges regardless of their proximity to talin nanoclusters (Fig. 5B) suggesting that other adaptors aside from talin might contribute to the activation of β1 by its engagement with the actin cytoskeleton. Indeed, dual color STED imaging showed co-enrichment of α5β1 nanoclusters (and the active β1 sub-population) together with tensin-3 at the FA periphery (Fig. S8). In contrast, the same analysis on active β3 nanoclusters shows a random distribution inside of FAs irrespective of their proximity to talin or other adaptor nanoclusters (Fig. 5C).

Active β1 nanoclusters form multiprotein hubs at the FA edges.

(A) Population of active β1-talin nanocluster pairs as a function of their distance-to-the-FA edge. The data plotted in light green correspond to active β1 nanoclusters located close to talin (i.e., iNND < median), while the inset dark green corresponds to active β1 nanoclusters located far from talin (i.e., iNND > median). iNND distributions are shown in Fig. S7B. Black lines are for simulated data sets. Bin width is 20 nm. (B) Variation from random for active β1-talin nanocluster pairs as function of their distance-to-the-FA-edge. Filled bars correspond to active β1 nanoclusters located close to talin (iNND < median) and empty bars correspond to active β1 nanoclusters located far from talin (iNND > median) (C) Similar to (B) but for active β3 nanoclusters. (D-F) Variation from random for paxillin (D), talin (E), vinculin (F) active β1 nanocluster pairs, as function of their distance-to-the-FA-edge, with data segregated in two different distributions according to the pairs proximity. Bin width is 20 nm.

Finally, we enquired whether adaptor protein nanoclusters (paxillin, talin, and vinculin) close to active β1 nanoclusters also locate closer to the FA edges. Remarkably, we found that all of these adaptor nanoclusters close to active β1 nanoclusters are also enriched on the FA edge region (∼150–200 nm from the edge), whereas adaptor nanoclusters far from active β1 appear to be more uniformly distributed (Fig. 5D–F). Taken together, our data reveal a differential mesoscale organization between active β1 and β3 nanoclusters inside FAs, with active β1 nanoclusters being enriched at the periphery of FAs alongside adhesion adaptor nanoclusters, while active β3 nanoclusters remain uniformly distributed throughout FAs, lacking peripheral enrichment.

Discussion

Here, we have combined dual-color super-resolution optical microscopy together with a robust set of analytical tools to provide insights on the lateral spatial organization of different adhesion proteins inside FAs. Using three different super-resolution imaging strategies, STORM, DNA-PAINT and STED we show that both α5β1 and αvβ3 integrins as well as their main adaptors, paxillin, talin and vinculin, form nanoclusters of similar size and molecular content inside FAs. We further validated our clustering analysis using Voronoi-tessellation in addition to DBSCAN rendering comparable results in terms of nanoclustering size and content. Moreover, this organization was independent on the membrane location, i.e., inside or outside FAs, or FA maturation, as cells inspected at different spreading times exhibited comparable nanoscale organization of these adhesion proteins.

Earlier reports showed that nascent adhesions (NAs) are formed due to integrin nanoclusters that either rapidly disassemble at the plasma membrane without force or further mature into FAs in response to mechanical forces (35, 59). Our results now show that even in the presence of mechanical forces (i.e., inside FAs), integrin nanoclusters remain segregated from each other while maintaining their sizes and number of molecules. Similar integrin nanocluster properties were also found outside adhesions, reinforcing the notion that integrin nanoclustering is force-independent (35). Interestingly, we found that the main cytoplasmic adaptors, paxillin, talin and vinculin, also form nanoclusters of similar size and number of single molecule localizations as the integrins, regardless of whether they locate inside or outside FAs. The presence of adhesion protein nanoclusters outside of FAs is intriguing. However, immobilized (active) integrins have been previously observed outside of FAs (25, 60). We suggest that these isolated nanoclusters may represent transient nascent adhesions forming on the cell membrane. Although super-resolution images do not allow us to directly extract the absolute number of proteins contained in nanoclusters given the inability to guarantee full labelling saturation and primary-to-secondary antibody stoichiometry, amongst others, the similarity in terms of number of single molecule localizations amongst integrins and their adaptor partners obtained from STORM and DNA-PAINT data suggest that these nanoclusters might be stoichiometrically similar. It will be interesting in the future to devise and/or to adapt emerging super-resolution strategies to determine the precise number of molecules contained in each nanocluster (61–63), although it might take a while until these approaches become routinely available to the community. Together, our results strongly indicate that adhesion proteins organize as universal nano-hubs, functioning as modular adhesion units to serve multiple integrin-dependent functions related to adhesion and migration but also to mechanosensing and mechanotransduction.

Consistent with previous results (36), our data revealed that active and inactive α5β1 nanoclusters coexist inside FAs but are fully segregated from each other. Remarkably, αvβ3 also exhibits similar nanoscale segregation, with total population-to-active integrin ratios indicating a similar segregation between active and inactive β3. Furthermore, earlier data from our group demonstrated that the integrin αLβ2 expressed on leukocytes also segregates into active and inactive nanoclusters (64, 65). Thus, lateral segregation between active and inactive integrin nanoclusters may represent a common feature to spatially coordinate integrin activation. How could such cooperativity in integrin nanocluster activation arise? Our data show a nearly 1:1 matching in the number of active integrin nanoclusters, and the main cytoplasmic adaptors talin and vinculin, known to activate integrins by their engagement to the actin cytoskeleton. The availability of a similar number of adaptor nanoclusters containing comparable number of molecules as the integrin nanoclusters suggests that inactive integrin nanoclusters are poised for their collective nanocluster activation, which is achieved upon the concurrent engagement of talin and vinculin nanoclusters (and most probably also paxillin). This mechanism has been shown to operate in NAs in which concurrent recruitment of talin and vinculin prior to force generation is required for integrin activation and further maturation of NAs (66). In addition, our data reveals the existence of a large percentage of inactive α5β1 and αvβ3 integrins inside FAs with little spatial overlap between integrins and main adaptors. As integrin activation is highly dynamic and transient, we propose that this pool of inactive integrin nanoclusters might be required to increase the probability of new engagements between integrins and their adaptors once the active integrin nanoclusters become disengaged, stabilizing as a whole the FAs for lifetimes much longer than the binding-unbinding dwell-times of integrins with their partners.

Quantification of super-resolution data showed a higher number of α5β1 nanoclusters compared to their αvβ3 counterparts inside FAs. The relative percentage of active β1 nanoclusters is also higher than that of αvβ3 and moreover these two integrin nanoclusters never intermix inside FAs. Interestingly, we found that while the nanocluster properties of α5β1 and αvβ3 integrins in terms of number of localizations and size were independent of FA maturation, their spatial distribution inside FAs was markedly different as a function of seeding time. Indeed, α5β1 nanocluster distribution was already established at 90 min, while αvβ3 nanocluster distribution appeared rather random at earlier seeding times and progressively organized reaching a well-defined lateral spacing at 24 hours of spreading time. As FA strengthening over time requires mechanical forces, our data strongly suggest that forces might play a differential role in the lateral distribution of both integrin nanoclusters over time. Further experiments lowering actomyosin contractility of the cells combined with the use of mechanical stretching devices or substrates of different stiffness would be required to further investigate this hypothesis.

Importantly, we found that active β1 nanoclusters preferentially locate at the periphery of FAs and in proximity with paxillin, talin and vinculin nanoclusters, while active β3 are more uniformly distributed inside FAs. The differential degree of α5β1 and αvβ3 in terms of expression, activation state, and lateral nano- and meso-scale segregation, strongly indicates that integrin activation and their distinct functions inside FAs are spatially regulated (Fig. 6). Based on our results we postulate that (i) α5β1, through the ring on the FA periphery and its close proximity to paxillin, talin and vinculin nanoclusters, primarily functions for substrate attachment in its high tensional state, engaging with actomyosin filaments; (ii) αvβ3 and α5β1 integrin nanoclusters in the central region of the FA undertake the role of mechanotransduction, being more dynamic in terms of activation and inactivation states and most probably engaged with shorter branched actin filaments; and (iii) the difference in these roles is controlled by the distinct mechanical forces exerted on the integrins, depending on their location inside FAs, and for α5β1 perhaps also on its different tensional states and/or degree of engagement with the actin cytoskeleton (67).

Spatial regulation and activation of integrin nanoclusters and protein adaptors inside focal adhesions.

α5β1, αvβ3, and main protein adaptors organize as small nanoclusters inside FAs physically segregated from each other. A considerable high fraction of α5β1 and αvβ3 integrins remain in an inactive state and do not intermix at the nanoscale. Active β1 nanoclusters (green) are preferentially enriched at the FA edges and reside in close proximity to the adaptor protein nanoclusters, possibly engaged to actomyosin filaments (brown). Integrin nanoclusters in the FA center are more randomly distributed with respect to one another, most probably engaged with branched cortical actin. According to our results we propose that α5β1 nanoclusters located at the FA periphery could be primarily involved in cell substrate attachment, whereas α5β1 and αvβ3 nanoclusters at the center of FAs engage in mechanosensing and mechanotransduction. For simplification, we show a cross-section of a FA where active β1 integrins at the edge are engaged with talin. However, our data also show the presence of active β1 nanoclusters spatially separated from talin and possibly engaged to other binding partners such as tensin-3 linking β1 to the actin cytoskeleton (not shown in the cartoon for the sake of clarity).

Intriguingly, our analysis showed that active β1 nanoclusters preferentially locate at the FA edges regardless of their proximity to talin nanoclusters (Fig. 5B). These data are striking since active integrins are expected to be bound to both its ligand and talin, and therefore we had anticipated the whole active β1 population to be in close proximity to talin nanoclusters. We can envisage two plausible explanations for active β1 to be far from talin. One possible scenario is the nature of talin itself. As a long rod-like protein, talin stretching experiments have shown that the N- and C- termini can be separated up to 250 nm due to actomyosin activity (68). Since in our experiments, we have so far detected talin by labeling it in its C-terminal domain, which is opposite to the integrin binding side, it is possible that active β1 is engaged to extended talin even for inter-cluster distances larger than the median iNND. Further experiments labeling talin at the closest proximity to the integrin binding site could address the influence of talin size and extension in relation to β1 nanocluster spatial proximity.

A second possibility is the existence of other adaptor protein(s), aside from talin, able to activate β1. Such a candidate might be tensin-3, which has been shown to activate integrins (in particular, if β1 is phosphorylated) by interacting with the actin cytoskeleton, providing mechanical coupling for integrin activation (69) . In addition, it has been recently shown that tensin might compete with talin to bind α5β1 during the formation of fAs from FAs (70). Interestingly, our dual color STED experiments labeling the total α5β1 or active β1 populations together with tensin-3, reveal a clear co-enrichment of these proteins at the FA periphery (Fig. S8), suggesting that active β1 nanoclusters at the periphery of FAs might engage to the actin cytoskeleton either by associating to talin and vinculin or to tensin-3. Why would α5β1 be preferentially located at the FA periphery and associating either with talin or tensin? Although we do not know yet what drives the preferential location of α5β1 nanoclusters to the FA periphery, it is known that KANK2 also exhibits enrichment at the FA periphery, regulates talin activity and it is involved in the formation of α5β1-enriched fAs (71). Therefore, it is highly probable that α5β1 enrichment at the FA periphery is a necessary step for their translocation from mature FAs to fAs to then assemble fibronectin into the fibrillar networks as found and needed in connective tissues. Consistent with this idea, we have observed similar α5β1 distribution on mouse embryonic fibroblasts, which are the main cells that produce fibrillar adhesions. Thus, α5β1 nanocluster distribution inside FAs might be physiologically important for the process of fibronectin fibrillogenesis. This hypothesis is further substantiated by our observations of tensin-3 and β1 co-enrichment at the FA periphery as it has been documented that tensin is important for fibronectin fibrillogenesis (72) and more recently, it has been shown that tensin-3 interaction with talin drives the formation of fibronectin-associated fAs (70). As such, α5β1 enrichment at the FA periphery could serve to efficiently translocate these integrins from FAs to fAs, probably in a talin-tensin dependent manner. Additional experiments including simultaneous super-resolution mapping of α5β1, talin and tensin3 in mature FAs will be required to fully validate our hypothesis.

In summary, in contrast to the classical textbook view where FAs are pictured as dense and homogenous assemblies of integrins fully engaged to their cytoplasmic adaptors, our work reveals a highly complex organization inside FAs, where regulation of different integrins and their activation state are highly coordinated in space. We believe that such spatial regulation provides the means for different integrins to synergize and/or diversify their functions in terms of adhesion, migration, mechanosensing and mechanotransduction. Earlier reports showed that proteins such as KANK1, KANK2, and β1 integrins are enriched at the edges of adhesions, with KANK2 reducing the force exerted by the cells through its binding to talin (71, 73). KANK proteins have also been implicated in enabling communication between microtubules and actin by binding to talin and liprin-β1 at FA edges (74). More recent data suggest that the localization of KANK1 around FAs is driven by liquid-liquid phase separation (75). Our findings together with those mentioned above suggest that the α5β1 nanocluster belt is an important feature of FAs and may be critical for their function and/or the translocation of these integrins to fAs. However, further research is needed to fully understand how the organization of these key proteins contributes to their roles in fAs. p130Cas- and FAK-dependent phase separation has been also suggested to promote integrin clustering and NA assembly (76). Whether phase separation also contributes to the preferential organization and activation of β1 at the FA edges as we reveal here is an intriguing topic that deserves further investigation.

Materials and Methods

Sample preparation and immunolabeling

HFF-1 cells (human foreskin fibroblasts) were obtained from ATCC and used at low passage number (passage <15). Cells were maintained in DMEM supplemented with 15% FBS and cultured in a humidified atmosphere at 37° C and 5% CO2. Experiments were performed using cells seeded on 8-well chambered cover glass slides (#1 Lab-Tek, Thermofisher, #155411) previously coated with fibronectin (FN) at full substrate coverage. FN coating was performed by incubating the wells with 10 µg/ml FN for 30 min at room temperature (RT). We note that FBS can contain vitronectin (VN) and other proteins that may act as ligands of integrins. However, since the glass slides are saturated with FN prior to incubation with cells and medium, FN remains the dominant ligand in our experiments. After seeding, cells were allowed to spread for 24 h at 37° C and 5% CO2, after which they were fixed with 4% PFA in PBS for 15 min at RT. Cells were permeabilized using 0.1% Triton X-100 in PBS for 10 min and blocked with 3% BSA in PBS for 40 min. Details of reagents used for cell culture and sample preparation are provided in Table S1. For STORM experiments, primary antibodies against total α5β1 (CD49e/ IIA1) and αvβ3 integrins (CD51/61), active β1 integrin (CD29- 9Eg7) and active β3 integrins (CD61-LIBS2), and the adaptor proteins paxillin (Y113), talin1 (actin-binding C-terminal epitope) and vinculin were diluted in blocking buffer and added to the cells for 30 min at RT. Details of the primary antibodies used across the different experimental approaches are provided in Table S2. After washing with PBS, secondary antibodies that were conjugated in-house with dye pairs Alexa Fluor 405-Alexa Fluor 647 (405-647 channel) or Cy3B-Alexa Fluor 647 (Cy3-647 channel) (see Table S3), were added to the cells for 30 min at RT. After washing with PBS, cells were stored in PBS until imaged.

Image acquisition

Single-molecule localization microscopy (SMLM) data were obtained on a commercial Nikon Eclipse Ti system (N-STORM), equipped with a 100x oil objective with NA 1.49 using TIRF illumination. The detector was an ANDOR technology EMCCD iXon 897 camera, with a 256x256 pixel ROI and a pixel size of 160 nm. The system has an Agilent technologies laser box with wavelengths of 405 nm, 560 nm and 647 nm. Details of the DNA-PAINT acquisition, performed on the same microscopy platform, are provided in the Supplementary Information section. For STORM experiments, imaging was performed in a buffer containing 100 mM of the reducing agent cysteamine mercaptoethylamine (MEA) (77 mg of MEA in 1 ml of 360 mM HCl, stock solution 1 M), the enzymatic oxygen scavenger Glox solution (0.5 mg/ml glucose oxidase, 40 μg/ml catalase), and 6.25% glucose in PBS. Details of the STORM imaging buffer reagents are found in Table S6. Dual color imaging was carried out cyclically and sequentially. In brief, we first detected the 405-647 channel in a 5-frame cycle consisting of 1 activator frame (405 nm laser on) followed by 4 reporter frames (647 nm laser on), followed by the Cy3-647 channel detection sequence, similarly composed of 1 activator frame (560 nm laser on) followed by 4 reporter frames (647 nm laser on). The frame rate was 50 Hz (20 ms per frame) for all frames. Imaging was carried out by cyclically repeating this 10-frame sequence until all the reporter’s dyes were exhausted (typically ∼70,000 frames).

Single molecule detection, localization precision, and SMLM image reconstruction

Single-molecule localization microscopy images were reconstructed using Insight3 software, provided by Bo Huang (UCSF, initially developed in Xiaowei Zhuang’s lab, Harvard University). In each imaging frame, individual molecules were identified using intensity thresholding (gray levels set between 600-60,000), then each intensity profile was fitted with a 2D Gaussian function to identify the central position, i.e., the coordinates that represent the spatial localization of that fluorescent molecule. The analysis was run over the entirety of each data set, resulting in a final output that is a list of x,y coordinates (localizations) corresponding to each individual molecule that has been identified. Insight3 software allows for drift correction using the correlated function of the image, for details see (77). For Dual-color STORM, the main source of crosstalk is the nonspecific activation of the Alexa 647 by the 647 nm laser (78), which is activator-independent and can occur in any frame. We implemented statistical crosstalk correction where the Insight3 software analyses the full image in 50x50 nm regions of interest (ROIs) and determines if each ROI is predominantly one color of the other. Based on the acquisition design (in our case, one activator frame followed by four reporter frames), the software determines the probability of detecting a nonspecific reporter molecule (78). To minimize color crosstalk, only those localizations detected in the second frame of each five-frame STORM imaging cycle (that is, the first of the four reporter frames) were considered for future analytical steps. Since each single molecule can experience numerous blinking events, the variation of the x,y coordinates from frame to frame determines the effective resolution (79). To measure this, we detected multiple localization events arising from individual molecules deposited on glass and calculated the standard deviation of the localization positions. We measured the effective resolution of our system to be ∼15 nm. These localizations detected from nonspecific antibodies on glass were subsequently used to set a threshold to define the minimum number of localizations required for a cluster to be included in the DBSCAN-based nanocluster analysis. We found that the antibodies on glass had a median number of localizations of 5 (for channel 1) and 3 (for channel 2). With the aim of discarding all localizations coming from nonspecific binding or binding to individual antibody-labelled proteins, we set the threshold at 10 localizations per nanocluster.

Generation of masks and classification of the localizations belonging to FAs, fAs or outside adhesions

A Fiji plugin, developed by the Lakadamyali lab, was implemented as an initial step to split the localization data from the STORM images into separate data sets corresponding to the localizations contained within defined ROIs (80). First, the data from the two channels of the reconstructed STORM image were imported and the localizations rendered together in a 1024x1024 pixels image (Fig. S1A). Next, we applied an intensity threshold, until the adhesion areas were visually detected, a Gaussian blurring of the image (by 2 pixels), and a second threshold filter until a smooth binary mask was obtained (Fig. S1B). The final mask retaining only the adhesion areas was obtained by manually selecting the regions corresponding to the adhesion complexes (identified by visual inspection based on their size and location). In the first steps, this Fiji plugin creates separate files containing the localizations that fall within the selected ROIs: all adhesions, glass, and outside adhesions (i.e., cell membrane that contain the remaining areas of the image that do not correspond to all adhesions or glass). In a second step the “all adhesions” file was again analyzed by the same plugin and all adhesions were split into FAs and fAs and rendered with Insight3. Fig. S1C-F shows the resulting rendered localizations for the different regions. These final two-color files for different cell regions were then individually and independently analyzed.

Image analysis

Clustering detection using the DBSCAN algorithm

DBSCAN is a density-based data clustering algorithm designed to be executed on arbitrary data sets with minimal input from the user (40). The algorithm works by identifying arbitrary-shaped clusters of data points using two user-defined input values: epsilon (ε), which is a radial distance that serves to define a local neighborhood from a randomly selected initial data point; and Nmin, which is the minimum number of points required to be within the initial point’s neighborhood, so it is considered part of the cluster seed. For our image analysis, spots on glass outside the cell regions were analyzed using ε = 20 nm and Nmin = 2, in order to define a minimum threshold for the number of localizations per nanocluster. Important to note that we carried out this analysis on glass areas, next to the cells, found in all of our STORM images. In this way we ensure having the same imaging conditions for in-cell nanoclusters and for spots-on-glass, however, at the expense of having a broader distribution of localization per spot-on-glass due to the presence of cell debris on the glass. When analyzing the nanoclusters found in ROIs within the cell, the DBSCAN parameters were set to ε = 20 nm and Nmin=3 localizations, which are common values used in the field (36, 45). Nevertheless, since multiple blinks of a fluorophore might result in artifactual clusters, we only considered true nanoclusters as those containing 10 or more localizations. Examples on the identification of different nanoclusters by DBSCAN are shown in Fig. S1G. The algorithm also provides the center of mass (CoM) coordinates of each nanocluster together with the number of localizations per nanocluster and their diameter.

Nearest neighbor distance (NND) analysis

To establish how the protein nanoclusters are distributed relative to one another, we used the MATLAB (MathWorks) function knnsearch. This function computes for each nanocluster the NND, which is the shortest distance from the CoM of a given nanocluster to its closest CoM nanocluster neighbor. We also used the same function to determine the iNND (i.e., NND between nanoclusters of different proteins). To measure the percentage of nanoclusters that overlap (colocalization fraction), we first computed the edge-to-edge distance between nearest neighbors (EED-NN) by subtracting from the NND the radii of the two-corresponding near-neighbor nanoclusters. From the EED-NN distributions we calculated the percentage of occurrences with EED-NN values below zero. The knnsearch function was also used to extract the distance-to-the-FA edge, i.e., the shortest distance between nanoclusters and the edge of adhesions. To that end, we computed the NND between the CoM of each nanocluster to the center coordinates of the list of pixels that fall on the edge of the FA mask.

Computational generation of uniform nanocluster distributions

To assess whether the distance distributions obtained from the experimental data (NND, EED-NN and distance-to-the-FA-edge), corresponded to a preferential type of organization or to a random spatial distribution, we performed simulations by generating in silico histograms of non-overlapping and randomly distributed nanoclusters inside FAs. For this, we first extracted the exact number and size of nanoclusters per adhesion from the experimental data. Our initial approach was to computationally generate the same number of nanoclusters and distribute them uniformly over the adhesion masks regardless of their physical size. However, this approach was not computationally suitable because, in many cases, large nanoclusters did not fit without overlap in the remaining space available within adhesions after randomly positioning the large majority of smaller nanoclusters. We therefore classified nanoclusters according to their physical size and positioned the largest nanoclusters first, followed by the remaining smaller ones. In addition, because our experimental data showed well-segregated nanoclusters, we imposed the condition that simulated nanoclusters should not overlap with one another. Fig. S9 shows the different steps followed to generate distributions of random nanoclusters inside FAs.

Statistical analysis

Statistical analysis was carried out using GraphPad Prism (version 9.12). The details of the individual statistical tests as well as the level of significance along with p-values are explicitly provided in the figure legends. No predetermined statistical method was employed for establishing the sample size, as the sample sizes were not selected based on a predefined effect size. In this context, multiple experiments were independently performed, employing various sample replicates, as detailed in Table S8.